---
title: "OAT School information"
author: "Stefanie Meliss"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

options(scipen = 999)
# empty work space
rm(list = ls())

# define directory
dir <- getwd()
dir <- gsub("/01_explore", "", dir)

# load libraries
library(kableExtra)
library(dplyr)
library(ggplot2)

# read in school_list info about OAT schools
school_list <- read.csv(file = file.path(dir, "misc", "OAT Linked establishments.csv"))

# schools must have been part of OAT for 4 or more years
cut_off_date <- as.Date("2021-09-01")
school_list$Joined.date <- as.Date(school_list$Joined.date, format =  "%d/%m/%Y")
school_list <- school_list[school_list$Joined.date <= cut_off_date, ]

# derive URNs
urn_list <- school_list$URN
```

### Linked establishments  
  
Records retrieved from [www.compare-school-performance.service.gov.uk](https://get-information-schools.service.gov.uk/Groups/Group/Details/4106?_ga=2.126526093.1025285281.1728492540-421756468.1727784748)  

```{r, echo = F, results='asis'}
kbl(school_list, caption = "OAT Linked establishments extract", row.names = F) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), fixed_thead = T)

```


### General school information (2022-2023)

Records retrieved from [www.compare-school-performance.service.gov.uk](https://www.compare-school-performance.service.gov.uk/download-data?download=true&regions=0&filters=GIAS&fileformat=csv&year=2022-2023&meta=false)  

```{r, echo = F, results='asis'}
# read in general school info
school_info <- read.csv(file = file.path(dir, "data", "performance-tables", "2022-2023", "2022-2023_england_school_information.csv"))

# read in meta data about school information
meta <- read.csv(file = file.path(dir, "data", "performance-tables", "2022-2023", "school_information_meta.csv"))
meta$Column <- NULL
meta$Explanation <- paste0(meta$Field.Name, " = ", meta$Description)

explanation <- paste(meta$Explanation, collapse = ", ")

# subset data to only include OAT schools
school_info <- school_info %>% filter(URN %in% urn_list)


note <- "Data from the following AP schools is not shown: "
schools <-  paste(school_list$Establishment.name[school_list$URN %in% setdiff(urn_list, school_info$URN)], collapse = ", ")

kbl(school_info, caption = "OAT 2022-2023 General school information from www.compare-school-performance.service.gov.uk", row.names = F) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), fixed_thead = T) %>%
  footnote(general = paste0(note, schools, "."),
           symbol = paste0(explanation, "."))

```

### School, pupils and their characteristics  
  
Records retrieved from [https://explore-education-statistics.service.gov.uk/find-statistics/school-pupils-and-their-characteristics](https://content.explore-education-statistics.service.gov.uk/api/releases/60f096b8-6ed9-4e9e-97ee-2ca83867d51e/files)  

```{r, echo = F, results='asis'}
# read in general school info
school_info <- read.csv(file = file.path(dir, "data", "school-pupils-and-their-characteristics_2023-24", "supporting-files", "spc_school_level_underlying_data.csv"))

# subset data to only include OAT schools
school_info <- school_info %>% filter(urn %in% urn_list)

# thin down columns
school_info <- school_info[, 1:grep("academy_flag", names(school_info))]

kbl(school_info, caption = "OAT 2023-2024 School characteristics from https://explore-education-statistics.service.gov.uk", row.names = F) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), fixed_thead = T) 

```

  
Variable name                                                                        |  Variable description
-----------------------------------------------------------------------------------  |  ---------------------------------------------------------------------------------
academy_flag                                                                         |  Academy indicator - Academy or otherwise
admissions_policy                                                                    |  School admissions policy - Selective or otherwise
denomination                                                                         |  Denomination - Religious denomination
female_boarders                                                                      |  female boarders
fte_pupils                                                                           |  FTE pupils
full_time_pupils                                                                     |  Full time pupils
headcount_of_pupils                                                                  |  Headcount - Total
headcount_total_female                                                               |  Headcount - female
headcount_total_male                                                                 |  Headcount - male
infants_taken_a_free_school_meal_on_census_day                                       |  Infants taken a free school meal on census day
male_boarders                                                                        |  male boarders
number_of_early_year_pupils_years_e1_and_e2                                          |  Number of early year pupils (years E1 and E2)
number_of_fsm_eligible_pupils_taking_a_free_school_meal_on_census_day                |  Number of FSM eligible pupils taking a free school meal on census day
number_of_key_stage_1_pupils_years_1_and_2                                           |  Number of key stage 1 pupils (years 1 and 2)
number_of_key_stage_2_pupils_years_3_to_6                                            |  Number of key stage 2 pupils (years 3 to 6)
number_of_key_stage_3_pupils_years_7_to_9                                            |  Number of key stage 3 pupils (years 7 to 9)
number_of_key_stage_4_pupils_years_10_and_11                                         |  Number of key stage 4 pupils (years 10 and 11)
number_of_key_stage_5_pupils_years_12_to_14                                          |  Number of key stage 5 pupils (years 12 to 14)
number_of_nursery_pupils_years_n1_and_n2                                             |  Number of nursery pupils (years N1 and N2)
number_of_pupils_known_to_be_eligible_for_free_school_meals                          |  Known to be eligible for free school meals
number_of_pupils_known_to_be_eligible_for_free_school_meals_performance_tables       |  Known to be eligible for free school meals (using performance tables methodology)
number_of_pupils_not_reception_or_key_stage_1_to_5                                   |  Number of pupils not reception or key stage 1 to 5
number_of_pupils_of_compulsory_school_age_and_above                                  |  Number of pupils of compulsory school age and above
number_of_pupils_used_for_fsm_calculation_in_performance_tables                      |  Number of pupils used for FSM calculation in Performance Tables
number_of_pupils_whose_first_language_is_known_or_believed_to_be_english             |  First language is known or believed to be English
number_of_pupils_whose_first_language_is_known_or_believed_to_be_other_than_english  |  First language is known or believed to be other than English
number_of_pupils_whose_first_language_is_unclassified                                |  First language is unclassified
number_of_reception_pupils_year_r                                                    |  Number of reception pupils (year R)
number_of_schools                                                                    |  Number of schools
part_time_pupils                                                                     |  Part time pupils
phase_type_grouping                                                                  |  School type - Filter by school type
sex_of_school_description                                                            |  Sex of school description - Mixed, boys, girls
total_boarders                                                                       |  Total boarders
type_of_establishment                                                                |  School type (detailed) - Filter by school type (further detail)
urban_rural                                                                          |  Urban or rural indicator - Urban or rural indicator


```{r, echo = F, results='asis'}
# read in general school info
n_pupils <- read.csv(file = file.path(dir, "data", "school-pupils-and-their-characteristics_2023-24", "supporting-files", "spc_school_level_underlying_data.csv"))

# subset data to only include OAT schools
n_pupils <- n_pupils %>% filter(urn %in% urn_list)

# change col names
names(n_pupils) <- gsub("X.", "percentage", names(n_pupils))

# thin down columns
n_pupils <- n_pupils[, !grepl("time.male|time.female", names(n_pupils))]
tmp <- n_pupils[, c("urn", "school_name")]
n_pupils <- n_pupils[, grep("school_size", names(n_pupils)):ncol(n_pupils)]
n_pupils <- merge(tmp, n_pupils, by = 0)
n_pupils$Row.names <- NULL

kbl(n_pupils, caption = "OAT 2023-2024 Pupils and their characteristics from https://explore-education-statistics.service.gov.uk", row.names = F) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), fixed_thead = T) 

```

```{r, echo = F, results='asis'}
# read in general school info
class_info <- read.csv(file = file.path(dir, "data", "school-pupils-and-their-characteristics_2023-24", "supporting-files", "spc_school_level_class_size_underlying_data.csv"))

# subset data to only include OAT schools
class_info <- class_info %>% filter(urn %in% urn_list)

# thin down columns
tmp <- class_info[, c("urn", "school_name")]

class_info <- class_info[, grep("number.of.classes.of.size.1.30.taught.by.one.teacher", names(class_info)):grep("average.size.of.one.teacher.classes", names(class_info))]

class_info <- merge(tmp, class_info, by = 0)
class_info$Row.names <- NULL

kbl(class_info, caption = "OAT 2023-2024 Class size characteristics from https://explore-education-statistics.service.gov.uk", row.names = F) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), fixed_thead = T) 

```

