---
title: "School capacity"
author: "Stefanie Meliss"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 4
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

options(scipen = 999)
# empty work space
rm(list = ls())

# define directory
dir <- getwd()
dir <- gsub("/01_explore", "", dir)

# load libraries
library(kableExtra)
library(dplyr)
library(ggplot2)

# source ambition theme
devtools::source_url("https://github.com/stefaniemeliss/ambition_theme/blob/main/ambition_theme.R?raw=TRUE")


# read in misc info about OAT schools
misc <- read.csv(file = file.path(dir, "misc", "OAT Linked establishments.csv"))
misc <- xlsx::read.xlsx(file = file.path(dir, "government_data.xlsx"), sheetName = "data_OAT")

# read in workforce data
df <- read.csv(file = file.path(dir, "data", "school-capacity_2022-23", "data", "school-capacity_200910-202223.csv"))
names(df) <- gsub("X...", "", names(df), fixed = T)

# read in info about dataset
dict <- read.csv(file = file.path(dir, "misc", "meta_schoolcapacity.csv"))
names(dict) <- gsub("X...", "", names(dict), fixed = T)
vars <- dict$Variable.name

# derive URNs
urn_list <- misc$urn[misc$phase_type_grouping == "State-funded secondary" & grepl("Urban", misc$urban_rural)]
# urn_list <- misc$urn

# derive years
years_list <- sort(unique(df$time_period))

# subset data to only include OAT schools
df <- df %>% filter(school_urn %in% urn_list)

# replace with NAs
df <- df %>%
  mutate(across(matches("y_capacity"), ~na_if(., "x"))) %>% # x = not available - information has not been collected or there are no estimates available at this level of aggregation.
  mutate(across(matches("y_capacity"), ~na_if(., "z"))) %>% # z = not applicable - statistic cannot be produced. For example where a denominator is not available to produce a percentage.
  mutate(across(matches("y_capacity"), ~na_if(., "c"))) %>% # c = confidential - where presentation of data would disclose confidential information
  mutate(across(matches("y_capacity"), ~na_if(., "u"))) # u = low reliability - values of the potentially low quality, for example where values of statistical significance have been calculated.

# make numeric
df[, grepl("y_capacity", names(df))] <- apply(df[, grepl("y_capacity", names(df))], MARGIN = 2, FUN = as.numeric)


# add number of years to df
df <- merge(df, df %>% group_by(school_urn) %>%
              summarise(total_years_of_data = n()), by = "school_urn")

# modify data for plotting
df$time_period <- factor(df$time_period)
df$school_urn <- factor(df$school_urn)

# create scaffold for difference in difference calculations
scaffold <- merge(data.frame(time_period = factor(years_list)), data.frame(school_urn = factor(urn_list)))
df <- merge(scaffold, df, by = c("school_urn", "time_period"), all.x = T)
df$Diff <- NA
df$Shift <- NA
```

School capacity  
  
Filename: school-capacity_200910-202223.csv  
Geographic levels: School  
Time period: 2009/10 to 2022/23  
Content summary: School places, which includes sixth form places and excludes nursery places, reported by local authorities. File also includes number of pupils on roll from the May School Census;  the number of pupils in places that exceed their school's capacity and the number of unfilled places.  
  
Variable names and descriptions for this file are provided below:  

```{r, results='asis', echo = F}
dict %>%
  kbl(row.names = F) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), fixed_thead = T) %>% 
  print()
```

Footnotes:

1. Includes mainstream state schools with capacity in any of the year groups from reception to year 11 on 1 May for the relevant academic year.
2. Primary places include all reported capacity in primary and middle deemed primary schools. Capacity excludes nursery places.
3. Secondary places include all reported capacity in secondary, middle-deemed secondary and all-through schools. Capacity includes sixth form places.
4. Number of pupils on roll for reception year group and above. Taken from the May school census for the relevant academic year, or gathered during the school capacity collection if census data was not available for a school.
5. 2019/20 data not available due to the cancellation of the 2020 School Capacity survey due to COVID-19.
6. Number of pupils in places that exceed their school's capacity is the difference between school places and number of pupils on roll, for schools where the number of pupils on roll is higher than the school’s reported capacity. Calculated at school level and then summed to national, regional or local authority level.




```{r working_chunk, echo=FALSE, eval=F}

if (trans == "RAW") {
  dep_var = var
  
} else if (trans == "LAGGED") {
  
  df <- df %>%
    group_by(school_urn) %>%
    mutate(Diff = get(var) - lag(get(var))) %>%
    as.data.frame()
  
  dep_var = "Diff"
  
} else if (trans == "2Y ROLLING") {
  
  df <- df %>%
    group_by(school_urn) %>%
    mutate(Shift = zoo::rollapply(get(var), width = 2, FUN = mean, align = 'right', fill = NA)) %>%
    as.data.frame()
  
  dep_var = "Shift"
  
}


# look at histogram by year
plt <- ggplot(data = df, aes(x = get(dep_var))) + 
  geom_histogram(bins = 25) + 
  facet_wrap(~get(group_var)) + 
  ambition_theme + 
  ylab("Count") +
  xlab(paste(y_lab, "-", trans))
print(plt)

# compute descriptives per year
out <- rbind(
  do.call("rbind",psych::describeBy(df[, dep_var], group = df[, group_var]))
)
# edit output
out$vars <- NULL
out <- round(out, 3)
# print output
kbl(out, caption = paste("Descriptives of variable", var, trans)) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), fixed_thead = T) %>%
  print()

# augmented Dickey-Fuller test
tseries::adf.test(na.omit(out$mean), k = 0) %>%
  print()

# plot overall time series
plt <- ggplot(data = df, aes(x = time_period, y = get(dep_var), group = school_urn, col = school_urn)) +
  geom_point() + 
  geom_line() + 
  ambition_theme + theme(legend.position = "none",  
                         axis.text.x = element_text(angle = 45, hjust = 1)) +
  ylab(paste(y_lab, "-", trans)) +
  xlab(x_lab)
print(plt)

# plot yearly average
out[, group_var] <- row.names(out)
plt <- ggplot(data = out, aes(x = time_period, y = mean)) +
  geom_point(size = 3) + 
  geom_errorbar(aes(ymin = mean-1.96*se, ymax = mean+1.96*se), width=.2,
                position=position_dodge(.9)) +
  geom_line(group = 1) + 
  ambition_theme + theme(legend.position = "none",  
                         axis.text.x = element_text(angle = 45, hjust = 1)) +
  ylab(paste(y_lab, "-", trans)) +
  xlab(x_lab)
print(plt)

# within school stats
tmp <- df %>% group_by(school_urn) %>%
  summarise(obs_per_school = sum(!is.na(get(dep_var))),
            within_school_mean = mean(get(dep_var), na.rm = T),
            within_school_sd = sd(get(dep_var), na.rm = T),
            snr = abs(mean(get(dep_var), na.rm = T))/sd(get(dep_var), na.rm = T),
            rsd = sd(get(dep_var), na.rm = T)/abs(mean(get(dep_var), na.rm = T)))
tmp[-1] %>% 
  filter(within_school_sd > 0 & within_school_mean != 0) %>%
  psych::describe(omit = T) %>%
  kbl(caption = paste("Descriptives of school-level", var, trans)) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), fixed_thead = T) %>% 
  print()

# show in boxplot
plt <- ggplot(data = tmp, aes(x = rsd )) +
  geom_boxplot() +
  geom_point(aes(y = 0), color = coral) +
  ambition_theme +
  theme(axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank()) +
  scale_x_continuous(labels = scales::percent) +
  xlab(paste("%RSD", y_lab, "- RAW")) # % relative standard deviation
print(plt)

```


```{r, echo = F, results='asis',fig.align='center', warning=FALSE}
# define DV and label
group_var = "time_period"
x_lab = "Academic years"

transforms <- c("RAW", "LAGGED", "2Y ROLLING")

for (var in vars) {
  
  y_lab = dict$Variable.description[dict$Variable.name == var]
  
  cat("\n## ", y_lab, "\n")
  cat("\n\n")
  
  for (trans in transforms) {
    
    cat("\n#### ", stringr::str_to_sentence(trans), "data\n")
    cat("\n\n")
    
    <<working_chunk>>
      
      cat("\n\n")
    
    
  }
  
}

```

