---
title: "Size of the school workforce"
author: "Stefanie Meliss"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 4
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

options(scipen = 999)
# empty work space
rm(list = ls())

# define directory
dir <- getwd()
dir <- gsub("/01_explore", "", dir)

# load libraries
library(kableExtra)
library(dplyr)
library(ggplot2)

# source ambition theme
devtools::source_url("https://github.com/stefaniemeliss/ambition_theme/blob/main/ambition_theme.R?raw=TRUE")


# read in misc info about OAT schools
misc <- read.csv(file = file.path(dir, "misc", "OAT Linked establishments.csv"))
misc <- xlsx::read.xlsx(file = file.path(dir, "government_data.xlsx"), sheetName = "data_OAT")

# read in workforce data
df <- read.csv(file = file.path(dir, "data", "school-workforce-in-england_2023", "data", "workforce_2010_2023_fte_hc_nat_reg_la_sch.csv"))
names(df) <- gsub("X...", "", names(df), fixed = T)

# read in info about dataset
dict <- read.csv(file = file.path(dir, "misc", "meta_workforce.csv"))

# derive URNs
urn_list <- misc$urn[misc$phase_type_grouping == "State-funded secondary" & grepl("Urban", misc$urban_rural)]
# urn_list <- misc$urn

# derive years
years_list <- sort(unique(df$time_period))

# subset data to only include OAT schools
df <- df %>% filter(school_urn %in% urn_list)

# replace with NAs
df <- df %>%
  mutate(across(matches("fte|hc|teach"), ~na_if(., "x"))) %>% # x = not available - information has not been collected or there are no estimates available at this level of aggregation.
  mutate(across(matches("fte|hc|teach"), ~na_if(., "z"))) %>% # z = not applicable - statistic cannot be produced. For example where a denominator is not available to produce a percentage.
  mutate(across(matches("fte|hc|teach"), ~na_if(., "c"))) %>% # c = confidential - where presentation of data would disclose confidential information
  mutate(across(matches("fte|hc|teach"), ~na_if(., "u"))) # u = low reliability - values of the potentially low quality, for example where values of statistical significance have been calculated.

# make numeric
df[, grepl("fte|hc|teach", names(df))] <- apply(df[, grepl("fte|hc|teach", names(df))], MARGIN = 2, FUN = as.numeric)

# add number of years to df
df <- merge(df, df %>% group_by(school_urn) %>%
              summarise(total_years_of_data = n()), by = "school_urn")

# modify data for plotting
df$time_period <- factor(df$time_period)
df$school_urn <- factor(df$school_urn)


# compute more vars
prefix <- c("fte", "fte_ft", "fte_pt", "hc", "hc_ft", "hc_pt")
for (x in prefix) {
  # data for teachers with qts
  df[, paste0(x, "_all_teachers_with_qts")] <- df[, paste0(x,  "_all_teachers")] - df[, paste0(x, "_all_teachers_without_qts")]
}

# create scaffold for difference in difference calculations
scaffold <- merge(data.frame(time_period = factor(years_list)), data.frame(school_urn = factor(urn_list)))
df <- merge(scaffold, df, by = c("school_urn", "time_period"), all.x = T)
df$Diff <- NA
df$Shift <- NA
```

Size of the school workforce - school level

Filename: workforce_2010_2023_fte_hc_nat_reg_la_sch.csv  
Geographic levels: School  
Time period: 2010/11 to 2023/24  
Content summary: Headcount and Full time Equivalent of staff by type of school and characteristics including post held, age group and ethnicity (additional file at school level)  

Variable names and descriptions for this file are provided below:


```{r, results='asis', echo = F}
dict[grepl("classroom", dict$Variable.name), ] %>%
  kbl(row.names = F) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), fixed_thead = T) %>% 
  print()
```

> [fte|hc]_all_teachers = [fte|hc]_classroom_teachers + [fte|hc]_leadership_teachers  
> [fte|hc]_all_teachers = [fte|hc]_ft_all_teachers + [fte|hc]_pt_all_teachers  
> [fte|hc]_all_teachers = [fte|hc]_all_teachers_without_qts + [fte|hc]_all_teachers_with_qts  


```{r working_chunk, echo=FALSE, eval=F}

if (trans == "RAW") {
  dep_var = var
  
} else if (trans == "LAGGED") {
  
  df <- df %>%
    group_by(school_urn) %>%
    mutate(Diff = get(var) - lag(get(var))) %>%
    as.data.frame()
  
  dep_var = "Diff"
  
} else if (trans == "2Y ROLLING") {
  
  df <- df %>%
    group_by(school_urn) %>%
    mutate(Shift = zoo::rollapply(get(var), width = 2, FUN = mean, align = 'right', fill = NA)) %>%
    as.data.frame()
  
  dep_var = "Shift"
  
}


# look at histogram by year
plt <- ggplot(data = df, aes(x = get(dep_var))) + 
  geom_histogram(bins = 25) + 
  facet_wrap(~get(group_var)) + 
  ambition_theme + 
  ylab("Count") +
  xlab(paste(y_lab, "-", trans))
print(plt)

# compute descriptives per year
out <- rbind(
  do.call("rbind",psych::describeBy(df[, dep_var], group = df[, group_var]))
)
# edit output
out$vars <- NULL
out <- round(out, 3)
# print output
kbl(out, caption = paste("Descriptives of variable", var, trans)) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), fixed_thead = T) %>%
  print()

# augmented Dickey-Fuller test
tseries::adf.test(na.omit(out$mean), k = 0) %>%
  print()

# plot overall time series
plt <- ggplot(data = df, aes(x = time_period, y = get(dep_var), group = school_urn, col = school_urn)) +
  geom_point() + 
  geom_line() + 
  ambition_theme + theme(legend.position = "none",  
                         axis.text.x = element_text(angle = 45, hjust = 1)) +
  ylab(paste(y_lab, "-", trans)) +
  xlab(x_lab)
print(plt)

# plot yearly average
out[, group_var] <- row.names(out)
plt <- ggplot(data = out, aes(x = time_period, y = mean)) +
  geom_point(size = 3) + 
  geom_errorbar(aes(ymin = mean-1.96*se, ymax = mean+1.96*se), width=.2,
                position=position_dodge(.9)) +
  geom_line(group = 1) + 
  ambition_theme + theme(legend.position = "none",  
                         axis.text.x = element_text(angle = 45, hjust = 1)) +
  ylab(paste(y_lab, "-", trans)) +
  xlab(x_lab)
print(plt)

# within school stats
tmp <- df %>% group_by(school_urn) %>%
  summarise(obs_per_school = sum(!is.na(get(dep_var))),
            within_school_mean = mean(get(dep_var), na.rm = T),
            within_school_sd = sd(get(dep_var), na.rm = T),
            snr = abs(mean(get(dep_var), na.rm = T))/sd(get(dep_var), na.rm = T),
            rsd = sd(get(dep_var), na.rm = T)/abs(mean(get(dep_var), na.rm = T)))
tmp[-1] %>% 
  filter(within_school_sd > 0 & within_school_mean != 0) %>%
  psych::describe(omit = T) %>%
  kbl(caption = paste("Descriptives of school-level", var, trans)) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), fixed_thead = T) %>% 
  print()

# show in boxplot
plt <- ggplot(data = tmp, aes(x = rsd )) +
  geom_boxplot() +
  geom_point(aes(y = 0), color = coral) +
  ambition_theme +
  theme(axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank()) +
  scale_x_continuous(labels = scales::percent) +
  xlab(paste("%RSD", y_lab, "- RAW")) # % relative standard deviation
print(plt)

```


```{r, echo = F, results='asis',fig.align='center', warning=FALSE}
vars <- names(df)[grepl("classroom|leadership|_all_teach", names(df))]
vars <- names(df)[grepl("classroom", names(df))]
vars <- vars[!grepl("non", vars)]

# define DV and label
group_var = "time_period"
x_lab = "Academic years"

transforms <- c("RAW", "LAGGED", "2Y ROLLING")

for (var in vars) {
  
  y_lab = dict$Variable.description[dict$Variable.name == var]
  
  cat("\n## ", y_lab, "\n")
  cat("\n\n")
  
  for (trans in transforms) {
    
    cat("\n#### ", stringr::str_to_sentence(trans), "data\n")
    cat("\n\n")
    
    <<working_chunk>>
      
      cat("\n\n")
    
    
  }
  
}

```

