---
title: "Donor pool filtering"
author: "Stefanie Meliss"
date: "`r Sys.Date()`"
output: html_document
---

Following initial SCMs computed using the *Synth* package, the definition of the donor pool is re-evaluated. This file summarises the effect of removing the region filter and/or urbanicity filter.  

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Set scientific notation penalty to avoid scientific notation in outputs
options(scipen = 999)

# Clear the workspace and run garbage collection
# rm(list = ls())
gc()

# Load necessary libraries
library(kableExtra)
library(dplyr)
library(data.table)

# Define directories based on the current working directory
dir <- getwd()
dir <- gsub("02_scm", "", dir)
dir <- normalizePath(dir, winslash = "/")
dir_data <- file.path(dir, "data")
dir_misc <- file.path(dir, "misc")

# Copy data files from the source directory to the target directory
file.copy(
  file.path(gsub("scm_feasibility", "edu_stats", dir_data), "data_pupils.csv"),
  dir_data,
  overwrite = TRUE
)
file.copy(
  file.path(gsub("scm_feasibility", "edu_stats", dir_data), "data_establishments_search.csv"),
  dir_data,
  overwrite = TRUE
)

# Load data from CSV files into data tables
swf <- fread(file = file.path(dir_data, "data_swf.csv"))
est <- fread(file = file.path(dir_data, "data_establishments_search.csv"), na.strings = "")

#### PROCESS DATA FOR ESTABLISHMENTS ####

# Get data for the treated school based on the ID
id_treated <- 3344650
est_treated <- est %>%
  filter(laestab == id_treated) %>%
  mutate(status = "treated") %>%
  as.data.frame()


```

## Initial filters

Donor pool equal to St Peters with respect to:  

- Phase of education  
- Region  
- Level of urbanicity (Urban or Rural)  
- Denomination  
- Completeness of observations on the outcome measure  
- Note: Boarding schools are excluded from donor pool

```{r, echo=FALSE}
# Get donor pool data excluding the treated school
est_cont <- est %>%
  filter(
    ! grepl("Boarding school", boarders_name),
    laestab != id_treated,
    phaseofeducation_name %in% unique(c(est_treated$phaseofeducation_name)),
    gor_name %in% unique(c(est_treated$gor_name)),
    urbanicity %in% c(unique(est_treated$urbanicity)), # coded as urban vs rural
    religiouscharacter_christian %in% unique(c(est_treated$religiouscharacter_christian))
  ) %>%
  mutate(status = "untreated") %>%
  as.data.frame()

# Save unique lists of laestab and urn
list_laestab <- c(unique(est_cont[, "laestab"]), unique(est_treated[, "laestab"]))
list_urn <- c(unique(est_cont[, "urn"]), unique(est_treated[, "urn"]))

#### CREATE OUTCOME DATASET ####

# Filter SWF data to create outcome dataset
z <- swf[laestab %in% list_laestab]
z[, status := fifelse(laestab %in% unique(est_treated[, "laestab"]), "treated", "untreated")]
z[, idx_treat := fifelse(laestab %in% unique(est_treated[, "laestab"]), TRUE, FALSE)]
z[, idx_donor := fifelse(laestab %in% unique(est_treated[, "laestab"]), FALSE, TRUE)]

# Remove rows with NA for the dependent variable and add observation count
var <- "pupil_to_qual_teacher_ratio"
z <- z %>%
  filter(!is.na(get(var))) %>%
  group_by(laestab) %>%
  mutate(
    obs_count = sum(!is.na(get(var)))
  ) %>%
  ungroup() %>%
  select(-one_of(c("region_code", "new_la_code", "old_la_code")))

# Filter for rows where observation count matches the treated ID and select columns
z <- z %>%
  filter(obs_count == unique(z$obs_count[z$laestab == id_treated])) %>%
  select(time_period, laestab, school, pupil_to_qual_teacher_ratio) %>%
  group_by(laestab) %>%
  arrange(laestab, desc(time_period)) %>%
  mutate(school = first(school)) %>%
  arrange(laestab, time_period) %>%
  ungroup() %>%
  as.data.frame()

# Update list of laestab numbers based on the dependent variable
list_laestab_dv <- unique(z$laestab)

cat("Schools in donor pool:", length(list_laestab_dv))
```

## Exclude region

Donor pool equal to St Peters with respect to:  

- Phase of education  
- Level of urbanicity (Urban or Rural)  
- Denomination  
- Completeness of observations on the outcome measure  
- Note: Boarding schools are excluded from donor pool

```{r, echo=FALSE}
# Get donor pool data excluding the treated school
est_cont <- est %>%
  filter(
    ! grepl("Boarding school", boarders_name),
    laestab != id_treated,
    phaseofeducation_name %in% unique(c(est_treated$phaseofeducation_name)),
    # gor_name %in% unique(c(est_treated$gor_name)),
    urbanicity %in% c(unique(est_treated$urbanicity)), # coded as urban vs rural
    religiouscharacter_christian %in% unique(c(est_treated$religiouscharacter_christian))
  ) %>%
  mutate(status = "untreated") %>%
  as.data.frame()

# Save unique lists of laestab and urn
list_laestab <- c(unique(est_cont[, "laestab"]), unique(est_treated[, "laestab"]))
list_urn <- c(unique(est_cont[, "urn"]), unique(est_treated[, "urn"]))

#### CREATE OUTCOME DATASET ####

# Filter SWF data to create outcome dataset
z <- swf[laestab %in% list_laestab]
z[, status := fifelse(laestab %in% unique(est_treated[, "laestab"]), "treated", "untreated")]
z[, idx_treat := fifelse(laestab %in% unique(est_treated[, "laestab"]), TRUE, FALSE)]
z[, idx_donor := fifelse(laestab %in% unique(est_treated[, "laestab"]), FALSE, TRUE)]

# Remove rows with NA for the dependent variable and add observation count
var <- "pupil_to_qual_teacher_ratio"
z <- z %>%
  filter(!is.na(get(var))) %>%
  group_by(laestab) %>%
  mutate(
    obs_count = sum(!is.na(get(var)))
  ) %>%
  ungroup() %>%
  select(-one_of(c("region_code", "new_la_code", "old_la_code")))

# Filter for rows where observation count matches the treated ID and select columns
z <- z %>%
  filter(obs_count == unique(z$obs_count[z$laestab == id_treated])) %>%
  select(time_period, laestab, school, pupil_to_qual_teacher_ratio) %>%
  group_by(laestab) %>%
  arrange(laestab, desc(time_period)) %>%
  mutate(school = first(school)) %>%
  arrange(laestab, time_period) %>%
  ungroup() %>%
  as.data.frame()

# Update list of laestab numbers based on the dependent variable
list_laestab_dv <- unique(z$laestab)

cat("Schools in donor pool:", length(list_laestab_dv))
```

## Exclude region but remove London

Donor pool equal to St Peters with respect to:  

- Phase of education  
- Level of urbanicity (Urban or Rural)  
- Denomination  
- Completeness of observations on the outcome measure  
- Note: Boarding schools are excluded from donor pool

```{r, echo=FALSE}
# Get donor pool data excluding the treated school
est_cont <- est %>%
  filter(
    ! grepl("Boarding school", boarders_name),
    laestab != id_treated,
    phaseofeducation_name %in% unique(c(est_treated$phaseofeducation_name)),
    ! grepl("London", gor_name),
    urbanicity %in% c(unique(est_treated$urbanicity)), # coded as urban vs rural
    religiouscharacter_christian %in% unique(c(est_treated$religiouscharacter_christian))
  ) %>%
  mutate(status = "untreated") %>%
  as.data.frame()

# Save unique lists of laestab and urn
list_laestab <- c(unique(est_cont[, "laestab"]), unique(est_treated[, "laestab"]))
list_urn <- c(unique(est_cont[, "urn"]), unique(est_treated[, "urn"]))

#### CREATE OUTCOME DATASET ####

# Filter SWF data to create outcome dataset
z <- swf[laestab %in% list_laestab]
z[, status := fifelse(laestab %in% unique(est_treated[, "laestab"]), "treated", "untreated")]
z[, idx_treat := fifelse(laestab %in% unique(est_treated[, "laestab"]), TRUE, FALSE)]
z[, idx_donor := fifelse(laestab %in% unique(est_treated[, "laestab"]), FALSE, TRUE)]

# Remove rows with NA for the dependent variable and add observation count
var <- "pupil_to_qual_teacher_ratio"
z <- z %>%
  filter(!is.na(get(var))) %>%
  group_by(laestab) %>%
  mutate(
    obs_count = sum(!is.na(get(var)))
  ) %>%
  ungroup() %>%
  select(-one_of(c("region_code", "new_la_code", "old_la_code")))

# Filter for rows where observation count matches the treated ID and select columns
z <- z %>%
  filter(obs_count == unique(z$obs_count[z$laestab == id_treated])) %>%
  select(time_period, laestab, school, pupil_to_qual_teacher_ratio) %>%
  group_by(laestab) %>%
  arrange(laestab, desc(time_period)) %>%
  mutate(school = first(school)) %>%
  arrange(laestab, time_period) %>%
  ungroup() %>%
  as.data.frame()

# Update list of laestab numbers based on the dependent variable
list_laestab_dv <- unique(z$laestab)

cat("Schools in donor pool:", length(list_laestab_dv))
```

## Exclude urbanicity

Donor pool equal to St Peters with respect to:  

- Phase of education  
- Region  
- Denomination  
- Completeness of observations on the outcome measure  
- Note: Boarding schools are excluded from donor pool

```{r, echo=FALSE}
# Get donor pool data excluding the treated school
est_cont <- est %>%
  filter(
    ! grepl("Boarding school", boarders_name),
    laestab != id_treated,
    phaseofeducation_name %in% unique(c(est_treated$phaseofeducation_name)),
    gor_name %in% unique(c(est_treated$gor_name)),
    # urbanicity %in% c(unique(est_treated$urbanicity)), # coded as urban vs rural
    religiouscharacter_christian %in% unique(c(est_treated$religiouscharacter_christian))
  ) %>%
  mutate(status = "untreated") %>%
  as.data.frame()

# Save unique lists of laestab and urn
list_laestab <- c(unique(est_cont[, "laestab"]), unique(est_treated[, "laestab"]))
list_urn <- c(unique(est_cont[, "urn"]), unique(est_treated[, "urn"]))

#### CREATE OUTCOME DATASET ####

# Filter SWF data to create outcome dataset
z <- swf[laestab %in% list_laestab]
z[, status := fifelse(laestab %in% unique(est_treated[, "laestab"]), "treated", "untreated")]
z[, idx_treat := fifelse(laestab %in% unique(est_treated[, "laestab"]), TRUE, FALSE)]
z[, idx_donor := fifelse(laestab %in% unique(est_treated[, "laestab"]), FALSE, TRUE)]

# Remove rows with NA for the dependent variable and add observation count
var <- "pupil_to_qual_teacher_ratio"
z <- z %>%
  filter(!is.na(get(var))) %>%
  group_by(laestab) %>%
  mutate(
    obs_count = sum(!is.na(get(var)))
  ) %>%
  ungroup() %>%
  select(-one_of(c("region_code", "new_la_code", "old_la_code")))

# Filter for rows where observation count matches the treated ID and select columns
z <- z %>%
  filter(obs_count == unique(z$obs_count[z$laestab == id_treated])) %>%
  select(time_period, laestab, school, pupil_to_qual_teacher_ratio) %>%
  group_by(laestab) %>%
  arrange(laestab, desc(time_period)) %>%
  mutate(school = first(school)) %>%
  arrange(laestab, time_period) %>%
  ungroup() %>%
  as.data.frame()

# Update list of laestab numbers based on the dependent variable
list_laestab_dv <- unique(z$laestab)

cat("Schools in donor pool:", length(list_laestab_dv))
```

## Exclude region and urbanicity

Donor pool equal to St Peters with respect to:  

- Phase of education  
- Denomination  
- Completeness of observations on the outcome measure  
- Note: Boarding schools are excluded from donor pool  

```{r, echo=FALSE}
# Get donor pool data excluding the treated school
est_cont <- est %>%
  filter(
    ! grepl("Boarding school", boarders_name),
    laestab != id_treated,
    phaseofeducation_name %in% unique(c(est_treated$phaseofeducation_name)),
    # gor_name %in% unique(c(est_treated$gor_name)),
    # urbanicity %in% c(unique(est_treated$urbanicity)), # coded as urban vs rural
    religiouscharacter_christian %in% unique(c(est_treated$religiouscharacter_christian))
  ) %>%
  mutate(status = "untreated") %>%
  as.data.frame()

# Save unique lists of laestab and urn
list_laestab <- c(unique(est_cont[, "laestab"]), unique(est_treated[, "laestab"]))
list_urn <- c(unique(est_cont[, "urn"]), unique(est_treated[, "urn"]))

#### CREATE OUTCOME DATASET ####

# Filter SWF data to create outcome dataset
z <- swf[laestab %in% list_laestab]
z[, status := fifelse(laestab %in% unique(est_treated[, "laestab"]), "treated", "untreated")]
z[, idx_treat := fifelse(laestab %in% unique(est_treated[, "laestab"]), TRUE, FALSE)]
z[, idx_donor := fifelse(laestab %in% unique(est_treated[, "laestab"]), FALSE, TRUE)]

# Remove rows with NA for the dependent variable and add observation count
var <- "pupil_to_qual_teacher_ratio"
z <- z %>%
  filter(!is.na(get(var))) %>%
  group_by(laestab) %>%
  mutate(
    obs_count = sum(!is.na(get(var)))
  ) %>%
  ungroup() %>%
  select(-one_of(c("region_code", "new_la_code", "old_la_code")))

# Filter for rows where observation count matches the treated ID and select columns
z <- z %>%
  filter(obs_count == unique(z$obs_count[z$laestab == id_treated])) %>%
  select(time_period, laestab, school, pupil_to_qual_teacher_ratio) %>%
  group_by(laestab) %>%
  arrange(laestab, desc(time_period)) %>%
  mutate(school = first(school)) %>%
  arrange(laestab, time_period) %>%
  ungroup() %>%
  as.data.frame()

# Update list of laestab numbers based on the dependent variable
list_laestab_dv <- unique(z$laestab)

cat("Schools in donor pool:", length(list_laestab_dv))
```

