---
title: "Results of cross-validated gridsearch with *scpi* package"
author: "Stefanie Meliss"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    self_contained: true
---
This markdown summarises the results of the grid search.  

The grid search iterated through the following parameters:  

- Was the outcome included as feature? (excl. DV vs. incl. DV)  
- Was any covariates adjusted for each feature? (none vs. Constant vs. Trend vs. Constant + Trend)  
- Which rolling window average width was applied to the features? (roll = 1 vs. roll = 2)  
- Was the rolling window also applied to the outcome variable? (Raw DV vs. Roll DV)
- Which years were included in the training data? (all years between 2010/11 and 2021/22)  
- Which schools were included in the donor pool? (Schools in the same region vs. schools in same and neighbouring region)  
- Was the donor pool filtered with respect to value of averaged pre-treatment outcome variable in relation to treated unit? (x SD thresholds)  
- Which weight constraints were defined during the estimation? (Simplex vs. L1-l2)  


Data up to 2021/22 was used to estimate the weights under various settings. The weights were validated using data from 2022/23 and 2023/24.  

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Clear the workspace and run garbage collection
rm(list = ls())
gc()

# load libraries
library(kableExtra)
library(dplyr)
library(data.table)
library(scpi)

# create function to source code
source_code <- function(root_dir_name = "code", target_repo = "helper_functions", branch = "main", file_name = "file.R") {
  
  # construct URL
  git_url <- paste0("https://raw.githubusercontent.com/stefaniemeliss/", target_repo, "/", branch, "/", file_name)
  
  # attempt to download from github
  tempp_file <- tempfile(fileext = ".R")
  message <- curl::curl_download(git_url, tempp_file, quiet = F)
  
  if(!grepl("Error", message)) {
    
    # if successful, source file
    source(tempp_file)
    remove(tempp_file)
    
  } else { # load local copy of file
    
    # Get the current working directory
    current_dir <- getwd()
    
    # Split the current directory into its components
    dir_components <- strsplit(current_dir, "/")[[1]]
    
    # Identify the root directory dynamically based on the provided root directory name
    root_index <- which(dir_components == root_dir_name)
    if (length(root_index) == 0) {
      stop(paste("Root directory", root_dir_name, "not found in the current path"))
    }
    root_dir <- do.call(file.path, as.list(dir_components[1:root_index]))
    
    # Identify the subdirectory one level below the root and construct its absolute path
    project_repo <- dir_components[root_index + 1]
    dir <- file.path(root_dir, project_repo)
    
    if (target_repo != project_repo) {
      dir <- gsub(project_repo, target_repo, dir) 
    }
    
    # Construct the full file path
    file_path <- file.path(dir, file_name)
    
    # Print the directory and file path for debugging
    print(paste("Directory:", dir))
    print(paste("File path:", file_path))
    
    # Source the file into the parent frame
    source(file_path, local = parent.frame())
  }
}

# source functions
source_code(target_repo = "scm_feasibility", file_name = "functions.R")

# Define the base directory
get_directory()

# get file stem name
file_stem <- get_file_stem()

# process data establishments #

# load in file with timeseries desc
summary <- read.csv(file.path(dir, "02_scm",  "02_treated_schools_filter_donor_pool_out.csv"))

# only select schools with sufficient donor pool
summary <- subset(summary, n_pool >= 50)

# save laestab numbers
list_laestab_treated <- unique(summary$laestab)
# list_laestab_treated <- list_laestab_treated[1:3] # debug
# list_laestab_treated <- list_laestab_treated[1:2] # debug

# create df_region as reference
df_region <- unique(summary[, c("laestab", "same", "neighbouring")])
```

```{r plot_ts, eval=FALSE, include=F}
#id_treated = list_laestab[1] # debug
rolling.window = 2
roll.vars <- c("pupil_to_qual_teacher_ratio", "fte_avg_age", "pnpupfsm_e")

# Define timeseries
years.avail <- sort(unique(df$time_period))
period.post <- c(2022:2023)
period.pre <- setdiff(years.avail, period.post)

# apply rolling window average to pre treatment period
pre <- df %>%
  filter(time_period %in% period.pre & laestab == id_treated) %>%
  arrange(desc(time_period)) %>%
  mutate(
    across(all_of(roll.vars), ~ zoo::rollapply(.x, width = rolling.window, mean, align = "left", partial = T), .names = "{col}_roll_cv"),
    fold = "Train"
  )

# apply rolling window average to pre treatment period
post <- df %>%
  filter(time_period %in% period.post & laestab == id_treated) %>%
  arrange(desc(time_period)) %>%
  mutate(
    across(all_of(roll.vars), ~ zoo::rollapply(.x, width = rolling.window, mean, align = "left", partial = T), .names = "{col}_roll_cv"),
    fold = "Validation"
  )

# combine rolled data from time time periods
df_treated <- bind_rows(pre, post) %>%
  arrange(desc(time_period)) %>%
  select(- pupil_to_qual_unqual_teacher_ratio) %>%
    mutate(
    across(all_of(roll.vars), ~ zoo::rollapply(.x, width = rolling.window, mean, align = "left", partial = T), .names = "{col}_roll")  
    ) %>%
    mutate(
    across(all_of(paste0(roll.vars, "_roll")), ~ ifelse(time_period > 2020, ., NA))  
    ) %>%
  as.data.frame()

lookup <- unique(df_treated[, c("time_period", "time_period_str")])


#### CREATE SPAGHETTI PLOT ####

# create data in long format
df_long <- df_treated %>%
  tidyr::pivot_longer(
    cols = -c(laestab, school, time_period, time_period_str, fold, school, urn),
    names_to = "variable") %>%
  mutate(category = case_match(variable, 
                               "pupil_to_qual_teacher_ratio" ~ "Outcome",
                               "pupil_to_qual_teacher_ratio_roll" ~ "Outcome",
                               "pupil_to_qual_teacher_ratio_roll_cv" ~ "Outcome",
                               "fte_avg_age" ~ "Teacher age",
                               "fte_avg_age_roll" ~ "Teacher age",
                               "fte_avg_age_roll_cv" ~ "Teacher age",
                               "pnpupfsm_e" ~ "% pupils FSM",
                               "pnpupfsm_e_roll" ~ "% pupils FSM",
                               "pnpupfsm_e_roll_cv" ~ "% pupils FSM"
  ),
  roll = case_match(variable, 
                    "pupil_to_qual_teacher_ratio" ~ "Raw",
                    "pupil_to_qual_teacher_ratio_roll" ~ "Rolling average",
                    "pupil_to_qual_teacher_ratio_roll_cv" ~ "Rolling average CV",
                    "fte_avg_age" ~ "Raw",
                    "fte_avg_age_roll" ~ "Rolling average",
                    "fte_avg_age_roll_cv" ~ "Rolling average CV",
                    "pnpupfsm_e" ~ "Raw",
                    "pnpupfsm_e_roll" ~ "Rolling average",
                    "pnpupfsm_e_roll_cv" ~ "Rolling average CV"
  )
  ) %>%
  as.data.frame()

df_long$category <- factor(df_long$category, levels = c("Outcome", "% pupils FSM", "Teacher age"))


# plot timeseries for each variable
plt <- ggplot(data = df_long, aes(x = time_period, y = value, col = roll, group = interaction(laestab, roll))) +
  geom_vline(xintercept = 2021.5, linetype = "dotted") +
  geom_line() + 
  geom_point() +
  facet_wrap(~ category, ncol = 1, strip.position = "top", scales = "free_y") +
  ambition_theme +
  ylab("Reported value") +
  xlab("Academic year") +
  scale_color_manual(
    breaks = c("Raw", "Rolling average CV", "Rolling average"),
    values = c("Raw" = red, "Rolling average CV" = blue, "Rolling average" = teal)) +
  scale_x_continuous(breaks = lookup$time_period, labels = lookup$time_period_str) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title.x = element_blank(), 
        legend.title = element_blank())

suppressWarnings(print(plt))
cat("\n\n")  # Add some space between plots


```

```{r grid, eval=FALSE, include=F}
#### process data grid search ####

# declare file name
file_name <- file.path(getwd(), paste0("03_treated_schools_scpi_cv_roll_gridsearch_", gsub(" ", "_", id_name), ".csv"))

# read in results
results <- read.csv(file_name)

# copy data for editing
tmp_results <- results
# apply(results[1:14], 2, unique)

tmp_results$rmspe_ratio <- tmp_results$rmspe_pre / tmp_results$rmspe_post

tmp_results[tmp_results$rmspe_pre < tmp_results$sd_treated & tmp_results$rmspe_post < tmp_results$sd_treated,  ]

# create new vars for facets #
# DV used
tmp_results$dv <- ifelse(grepl("pupil_to_qual_teacher_ratio", tmp_results$outcome.var), "Qualified", "All")
# is DV in features?
tmp_results$feat_dv <- ifelse(grepl("ratio", tmp_results$features), "incl. DV", "excl. DV")
tmp_results$feat_dv <- factor(tmp_results$feat_dv)
# covariate adjustment
tmp_results$adj <- ifelse(is.na(tmp_results$cov.adj), "none", tmp_results$cov.adj)
tmp_results$adj <- factor(tmp_results$adj, levels = c("none", "constant", "trend", "constant, trend"),
                          labels = c("none", "Constant", "Trend", "C + T"))
# which region was used for filtering
tmp_results$region <- ifelse(grepl("neighbouring", tmp_results$region.filter), "Same + Neighb.", "Same")
tmp_results$region <- factor(tmp_results$region)
# which SD filter was applied?
if ("50" %in% unique(tmp_results$sd.range)) {
  # Primary
  tmp_results$sd <- ifelse(grepl("50", tmp_results$sd.range), "0.5 SD", "1 SD")
} else {
  # Secondary
  tmp_results$sd <- ifelse(grepl("100", tmp_results$sd.range), "1 SD", "No SD")
}
tmp_results$sd <- factor(tmp_results$sd, levels = sort(unique(tmp_results$sd), decreasing = T))
# was rolling average computed
tmp_results$roll_w <- ifelse(tmp_results$rolling.window == 1, "Raw IV", "Roll IV")
tmp_results$roll_w <- factor(tmp_results$roll_w)
# was rolling average computed for outcome
tmp_results$roll_out <- ifelse(tmp_results$roll.outcome, "Roll DV", "Raw DV")
tmp_results$roll_out <- factor(tmp_results$roll_out)
# which years were used in training
tmp_results$pretreat <- ifelse(grepl("2019, 2020, 2021", tmp_results$period.pre), "all years", 
                               ifelse(!grepl("2019, 2021", tmp_results$period.pre), "excl. 20/21",
                                      "excl. 20/21 & 21/22"))
tmp_results$pretreat <- factor(tmp_results$pretreat)
# which constraints were applied to the weights
tmp_results$method <- tstrsplit(tmp_results$w.constr, "; ")[[1]]
tmp_results$method <- gsub("name = ", "", tmp_results$method)
tmp_results$method <- factor(tmp_results$method, levels = c("simplex", "L1-L2", "lasso", "user provided"))

# finalise dataset
tmp_results <- tmp_results %>%
  # reorder dataset
  arrange(feat_dv, adj, roll_out, roll_w, pretreat, region, sd, method) %>%
  # create id col
  mutate(it = as.character(1:nrow(tmp_results))) %>%
  # select columns
  select(c(it, dv, feat_dv, adj, region, sd, roll_out, roll_w, pretreat, method, sd_treated, rmspe_pre, rmspe_post))


# convert to long format
tmp_long <- tmp_results %>%
  tidyr::pivot_longer(
    cols = c(rmspe_pre, rmspe_post),
    names_to = "data",
    values_to = "rmspe") %>%
  mutate(data = case_match(data, 
                           "rmspe_pre" ~ "T",
                           "rmspe_post" ~ "V")) %>%
  mutate(data = factor(data)) %>%
  # Explicitly arrange dataset by all relevant variables
  arrange(feat_dv, adj, roll_out, roll_w, pretreat, region, sd, method, data)

# Define explicit numeric dodge positions
method_levels <- unique(tmp_results$method)
dodge_width <- 0.6
n_methods <- length(method_levels)
# Create evenly spaced numeric offsets around 0
method_positions <- seq(-dodge_width/2, dodge_width/2, length.out = n_methods)

# Create a lookup table
position_lookup <- tibble(
  method = method_levels,
  method_position = method_positions
) %>%
  mutate(method = factor(method, levels = c("simplex", "L1-L2", "lasso", "user provided"))
  )

# Merge these positions into your data
tmp_long <- tmp_long %>%
  left_join(position_lookup, by = "method") %>%
  mutate(
    data_numeric = as.numeric(factor(data)), # numeric x-axis position per 'data'
    x_position = data_numeric + method_position # final numeric position for plotting
  )

#### plot data grid search ####

plt <- ggplot(data = tmp_long, aes(x = x_position, y = rmspe, group = it)) +
  geom_hline(data = function(y) y %>% group_by(feat_dv, adj, roll_out, roll_w, pretreat, region, sd) %>% summarise(sd_treated = unique(sd_treated), .groups = 'drop'),
             aes(yintercept = sd_treated),color = black40, linetype = "dashed") +
  geom_point(aes(colour = method), size = 1) +
  geom_line(aes(colour = method)) +
  scale_x_continuous(
    breaks = unique(tmp_long$data_numeric),
    labels = unique(tmp_long$data)
  ) +
  ggh4x::facet_grid2(cols = vars(feat_dv, adj, roll_out, roll_w), rows = vars(pretreat, region, sd), strip = ggh4x::strip_nested()) +
  ambition_theme +
  theme(strip.text = element_text(size = 10),
        legend.title = element_text(size = 10),
        legend.text = element_text(size = 10),
        axis.title = element_text(size = 10),
        axis.text = element_text(size = 10),
        #axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        plot.caption = element_text(size = 10),
        legend.margin=margin(t = -8)) +
  ylab("Root Mean Squared Prediction Error (RMSPE)") +
  xlab("Cross-validation fold") +
  guides(colour = guide_legend(title = "Weight constraint")) +
  scale_colour_manual(values = ambition_palette_bright)
print(plt)
cat("\n\n")


```

```{r, echo=F, message=FALSE, results='asis', fig.align='center', fig.height=10, fig.width=10}
for (i in 1:length(list_laestab_treated)) {
  
  id_treated <- list_laestab_treated[i]
  
  # process data
  df <- process_data_scm(id_treated = id_treated, read_files = T, export_data.tables = T)
  regions <- est_treated$gor_name
  
  # determine ids of control schools
  id_cont <- unique(df$laestab[df$laestab != id_treated])
  
  cat("# ", id_name, "\n\n")
  
  # print some information about the school
  cat("Type of establishment:", est_treated$typeofestablishment_name, "\n\n")
  cat("Phase:", est_treated$phaseofeducation_name, "\n\n")
  cat("Gender of pupils:", est_treated$gender_name, "\n\n")
  cat("Religious character:", est_treated$religiouscharacter_name, "\n\n")
  cat("Trust flag:", est_treated$trustschoolflag_name, "\n\n")
  cat("Local authority:", est_treated$la_name, "\n\n")
  cat("Region:", est_treated$gor_name, "\n\n")
  
  cat("### Timeseries plots \n\n")
  # Plot data
  <<plot_ts>>
    
    # run grid searcg
    cat("### Grid search \n\n")
  <<grid>>
    
}


```
