---
title: "AugSynth"
author: "Stefanie Meliss"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
---

Source: https://github.com/ebenmichael/augsynth/tree/master/vignettes


```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
  )
library(kableExtra)
```

# `augsynth`: The Augmented Synthetic Control Method


## Installation

You can install `augsynth` from github using `devtools`.

```{r install, results="hide", message=F, eval=F}
## Install devtools if noy already installed
install.packages("devtools", repos='http://cran.us.r-project.org')
## Install augsynth from github
devtools::install_github("ebenmichael/augsynth")
```

## Example: Effects of the 2012 Kansas Tax Cuts 

### The data
To show the usage and features of `augsynth`, we'll use data on the impact of personal income tax cuts in Kansas that comes with the `AugSynth` package. Our interest is in estimating the effect of income tax cuts on gross state product (GSP) per capita.

```{r load_data, results="hide", warning=F, message=F}
library(magrittr)
library(dplyr)
library(augsynth)
data(kansas)
```

The `kansas` dataset contains the GSP per capita (the outcome measure) `lngdpcapita` for all 50 states from the first quarter of 1990 to the first quarter of 2016.

To run `augsynth`, we need to include a treatment status column that indicates which region was treated and at what time. The table in `kansas` contains the column `treated` to denote this. In the original study, the second quarter of 2012 was the implementation of the tax cut in Kansas.

```{r treated_units}
kansas %>% 
  select(year, qtr, year_qtr, state, treated, gdp, lngdpcapita) %>% 
  filter(state == "Kansas" & year_qtr >= 2012 & year_qtr < 2013) 
```


### Synth
Now to find a synthetic control using the entire series of pre-intervention outcomes (and no auxiliary covariates), we can use `augsynth`. To do so we just need to give `augsynth` a formula like `outcome ~ treatment`, tell it what the unit and time variables are, optionally provide when intervention took place (the code will automatically determine this if `t_int` is not provided), and specify that we don't want to fit an outcome model

```{r fit_synth, message=F, warning=F}
library(augsynth)
syn <- augsynth(lngdpcapita ~ treated, fips, year_qtr, kansas,
                progfunc = "None", scm = T)
```

We can then look at the ATT estimates for each post-intervention time period and overall. 
We'll also see the quality of the synthetic control fit measured by the L2 distance between Kansas and its synthetic control, and the percent improvement over uniform weights.
By default, we'll also see pointwise confidence intervals using a [conformal inference procedure](https://arxiv.org/abs/1712.09089).

```{r summ_syn}
summary(syn)
```


The default test statistic is the sum of the absolute treatment efects `function(x) sum(abs(x))`. We can change the test statistic via the `stat_func` argument. For instance, if we want to perform a one-way test against postive effects, we can set the test stastic to be the negative sum `function(x) -sum(x)`:
```{r summ_syn_neg}
summary(syn, stat_func = function(x) -sum(x))
```
Or if we want to priotize testing the average post-treatment effect, we can set it to be the absolute sum:
```{r summ_syn_sum}
summary(syn, stat_func = function(x) abs(sum(x)))
```


It's easier to see this information visually. Below we plot the difference between Kansas and it's synthetic control. Before the tax cuts (to the left of the dashed line) we expect these to be close, and after the tax cuts we measure the effect (with point-wise confidence intervals).

```{r fig_syn, fig.width=8, fig.height=4.5, echo=T, fig.align="center"}
plot(syn)
```

We can also compute point-wise confidence intervals using the [Jackknife+ procedure](https://arxiv.org/abs/1905.02928) by changing the `inf_type` argument, although this requires additional assumptions.

```{r fig_syn_plus, fig.width=8, fig.height=4.5, echo=T, fig.align="center"}
plot(syn, inf_type = "jackknife+")
```


### Augmenting synth with an outcome model
In this example the pre-intervention synthetic control fit has an L2 imbalance of 0.083, about 20% of the imbalance between Kansas and the average of the other states. We can reduce this by _augmenting_ synth with ridge regression. To do this we change `progfunc` to `"Ridge"`. We can also choose the ridge hyper-parameter by setting `lambda`, while not specifying `lambda` will determine one through cross validation:
```{r fit_asynth, message=F, warning=F}
asyn <- augsynth(lngdpcapita ~ treated, fips, year_qtr, kansas,
                progfunc = "Ridge", scm = T)
```

We can plot the cross-validation MSE when dropping pre-treatment time periods by setting `cv = T` in the `plot` function:

```{r fig_asyn_cv, fig.width=8, fig.height=4.5, echo=T, fig.align="center"}
plot(asyn, cv = T)
```

By default, the CV procedure chooses the maximal value of `lambda` with MSE within one standard deviation of the minimal MSE. To instead choose the `lambda` that minizes the cross validation MSE, set `min_1se = FALSE`.


We can look at the summary and plot the results. Now in the summary output we see an estimate of the overall bias of synth; we measure this with the average amount that augmentation changes the synth estimate. Notice that the estimates become somewhat larger in magnitude, and the standard errors are tighter.
```{r summ_asyn}
summary(asyn)
```

```{r fig_asyn, fig.width=8, fig.height=4.5, echo=T, fig.align="center"}
plot(asyn)
```

There are also several auxiliary covariates. We can include these in the augmentation by fitting an outcome model using the auxiliary covariates. To do this we simply add the covariates into the formula after `|`. By default this will create time invariant covariates by averaging the auxiliary covariates over the pre-intervention period, dropping `NA` values. We can use a custom aggregation function by setting the `cov_agg` argument. Then the lagged outcomes and the auxiliary covariates are jointly balanced by SCM and the ridge outcome model includes both.

```{r fit_covsynth, message=F, warning=F}
covsyn <- augsynth(lngdpcapita ~ treated | lngdpcapita + log(revstatecapita) +
                                           log(revlocalcapita) + log(avgwklywagecapita) +
                                           estabscapita + emplvlcapita,
                   fips, year_qtr, kansas,
                   progfunc = "ridge", scm = T)

```

Again we can look at the summary and plot the results.
```{r summ_cvsyn}
summary(covsyn)
```

```{r fig_covsyn, fig.width=8, fig.height=4.5, echo=T, fig.align="center"}
plot(covsyn)
```

Now we can additionally fit ridge ASCM on the residuals, look at the summary, and plot the results.
```{r fit_covsynth_aug, message=F, warning=F}

covsyn_resid <- augsynth(lngdpcapita ~ treated | lngdpcapita + log(revstatecapita) +
                                           log(revlocalcapita) + log(avgwklywagecapita) +
                                           estabscapita + emplvlcapita,
                   fips, year_qtr, kansas,
                   progfunc = "ridge", scm = T, lambda = asyn$lambda,
                   residualize = T)
```

```{r summ_cvsyn_resid}
summary(covsyn_resid)
```


```{r fig_covsyn_resid, fig.width=8, fig.height=4.5, echo=T, fig.align="center"}
plot(covsyn_resid)
```


Finally, we can augment synth with many different outcome models. The simplest outcome model is a unit fixed effect model, which we can include by setting `fixedeff = T`.
```{r fit_desyn, message=F, warning=F}

desyn <- augsynth(lngdpcapita ~ treated,
                   fips, year_qtr, kansas,
                   progfunc = "none", scm = T,
                   fixedeff = T)
```


```{r summ_desyn}
summary(desyn)
```


```{r fig_desyn, fig.width=8, fig.height=4.5, echo=T, fig.align="center"}
plot(desyn)
```

We can incorproate other outcome models by changing the `progfunc`.
Several outcome models are available, including, fitting the factor model directly with `gsynth`, general elastic net regression, bayesian structural time series estimation with `CausalImpact`, and matrix completion with `MCPanel`. For each outcome model you can supply an optional set of parameters, see documentation for details.




# `augsynth`: Estimating treatment effects with staggered adoption

### The data

To show the features of the `multisynth` function we will use data on the effects of states implementing mandatory collective bargaining agreements for public sector unions [(Paglayan, 2018)](https://onlinelibrary.wiley.com/doi/full/10.1111/ajps.12388)

```{r results="hide", warning=F, message=F}
library(magrittr)
library(dplyr)
library(augsynth)
```

```{r }
data <- read.csv("https://dataverse.harvard.edu/api/access/datafile/:persistentId?persistentId=doi:10.7910/DVN/WGWMAV/3UHTLP", sep="\t")
```

The dataset contains several important variables that we'll use:

- `year`, `State`: The state and year of the measurement
- `YearCBrequired`: The year that the state adopted mandatory collective bargaining
- `lnppexpend`: Log per pupil expenditures in constant 2010 $

```{r echo = F}
data %>% 
    filter(year == 1960) %>% 
    select(year, State, YearCBrequired, lnppexpend) %>%
    head() %>%
    kable() %>%
    kable_styling(bootstrap_options =c("hover", "responsive"))
```

To run `multisynth`, we need to include a treatment status column that indicates which state is treated in a given year, we call this `cbr` below. We also restrict to the years 1959-1997 where we have yearly measurements of expenditures and drop Washington D.C. and Wisconsin from the analysis.

```{r }
data %>%
    filter(!State %in% c("DC", "WI"),
           year >= 1959, year <= 1997) %>%
    mutate(YearCBrequired = ifelse(is.na(YearCBrequired), 
                                   Inf, YearCBrequired),
           cbr = 1 * (year >= YearCBrequired)) -> analysis_df
```

## Partially pooled SCM with an intercept

To fit partially pooled synthetic controls, we need to give `multisynth` a formula of the form `outcome ~ treatment`, point it to the unit and time variables, and choose the level of partial pooling `nu`. Setting `nu = 0` fits a separate synthetic control for each treated unit and setting `nu = 1` fits fully pooled synthetic controls. If we don't set `nu`, `multisynth` will choose a heuristic value based on how well separate synthetic controls balance the overall average.
By default, `multisynth` includes an intercept shift along with the weights; we can exclude the intercept shift by setting `fixedeff = F`.
We can also set the number of pre-treatment time periods (lags) that we want to balance with the `n_lags` argument and the number of post-treatment time periods (leads) that we want to estimate with the `n_leads` argument. By default `multisynth` sets `n_lags` and `n_leads` to the number of pre-treatment and post-treatment periods for the last treated unit, respectively.

```{r }
# with a choice of nu
ppool_syn <- multisynth(lnppexpend ~ cbr, State, year, 
                        nu = 0.5, analysis_df)
# with default nu
ppool_syn <- multisynth(lnppexpend ~ cbr, State, year, 
                        analysis_df)

print(ppool_syn$nu)

ppool_syn
```

Using the `summary` function, we'll compute the treatment effects and standard errors and confidence intervals for all treated units as well as the average via the wild bootstrap. (This takes a bit of time so we'll store the output) We can also change the significant level associated with the confidence intervals by setting the `alpha` argument, by default `alpha = 0.05`.

```{r}
ppool_syn_summ <- summary(ppool_syn)
```

We can then report the level of global and individual balance as well as estimates for the average.

```{r }
ppool_syn_summ
```

`nopool_syn_summ$att` is a dataframe that contains all of the point estimates, standard errors, and lower/upper confidence limits. `Time = NA` denotes the effect averaged across the post treatment periods.

```{r echo = F}
ppool_syn_summ$att %>%
  filter(Time >= 0) %>%
  head() %>%
  kable() %>%
  kable_styling(bootstrap_options =c("hover", "responsive"))
```

We can also visually display both the pre-treatment balance and the estimated treatment effects.

```{r ppool_syn_plot, fig.width=8, fig.height=4.5, fig.align="center", warning=F, message=F}
plot(ppool_syn_summ)
```

And again we can hone in on the average effects.

```{r ppool_syn_plot_avg, fig.width=8, fig.height=4.5, fig.align="center", warning=F, message=F}
plot(ppool_syn_summ, levels = "Average")
```


### Collapsing into time cohorts

We can also collapse treated units with the same treatment time into _time cohorts_, and find one synthetic control per time cohort by setting `time_cohort = TRUE`. When the number of distinct treatment times is much smaller than the number of treated units, this will run significantly faster.

```{r }
# with default nu
ppool_syn_time <- multisynth(lnppexpend ~ cbr, State, year,
                        analysis_df, time_cohort = TRUE)

print(ppool_syn_time$nu)

ppool_syn_time
```

We can then compute effects for the overall average as well as for each treatment time cohort, rather than individual units.

```{r}
ppool_syn_time_summ <- summary(ppool_syn_time)
ppool_syn_time_summ
```

```{r echo = F}
ppool_syn_time_summ$att %>%
  filter(Time >= 0) %>%
  head() %>%
  kable() %>%
  kable_styling(bootstrap_options =c("hover", "responsive"))
```

Again we can plot the effects.

```{r ppool_syn_time_plot, fig.width=8, fig.height=4.5, fig.align="center", warning=F, message=F}
plot(ppool_syn_time_summ)
```


### Including auxiliary covariates

We can also include an additional set of covariates to balance along with the pre-treatment outcomes. First, let's create a data frame with the values of some covariates in a few different years:

```{r cov_data}

data %>%
  select(State, year, agr, pnwht, purban, perinc, studteachratio) %>%
  group_by(State) %>%
  summarise(perinc_1959 = perinc[year == 1959],
            studteachratio_1959 = studteachratio[year == 1959]) %>% 
  # filter to lower 48 where we have data
  filter(!State %in% c("AK", "HI"))  -> cov_data

analysis_df %>%
  inner_join(cov_data, by = "State") -> analysis_df_covs

```

To include auxiliary covariates, we can add them in to the formula after `|`. This will balance the auxiliary covariates along with the pre-treatment outcomes simultanouesly. If the covariates vary during the pre-treatment periods, `multisynth` will use the average pre-treatment value. We can change this behavior by including our own custom aggregation function via the `cov_agg` argument.
```{r cov_syn}
# with default nu
ppool_syn_cov <- multisynth(lnppexpend ~ cbr | perinc_1959 + studteachratio_1959,
                            State, year, analysis_df_covs)

print(ppool_syn_cov$nu)

ppool_syn_cov
```

Again we can compute effects, along with their standard errors and confidence intervals, and plot.
```{r}
ppool_syn_cov_summ <- summary(ppool_syn_cov)
ppool_syn_cov_summ
```

```{r echo = F}
ppool_syn_cov_summ$att %>%
  filter(Time >= 0) %>%
  head() %>%
  kable() %>%
  kable_styling(bootstrap_options =c("hover", "responsive"))
```

Again we can plot the effects.
```{r ppool_syn_cov_plot, fig.width=8, fig.height=4.5, fig.align="center", warning=F, message=F}
plot(ppool_syn_cov_summ, levels = "Average")
```

# `augsynth`: Estimating multiple outcome effects

### The data
To demonstrate `augsynth` with multiple outcomes, we'll use data on the impact of personal income tax cuts in Kansas that comes with the `AugSynth` package. Our interest is in estimating the effect of income tax cuts on gross state product (GSP) per capita, wages, establishment counts, and other macroeconomic indicators.

```{r load_data2, results="hide", warning=F, message=F}
library(magrittr)
library(dplyr)
library(augsynth)
data(kansas)
```

The `kansas` dataset contains the GSP per capita (the outcome measure) `lngdpcapita` for all 50 states from the first quarter of 1990 to the first quarter of 2016.

To run `augsynth`, we need to include a treatment status column that indicates which region was treated and at what time. The table in `kansas` contains the column `treated` to denote this. In the original study, the second quarter of 2012 was the implementation of the tax cut in Kansas.

```{r treated_units2}
kansas %>% select(year, qtr, year_qtr, state, treated, gdp, lngdpcapita) %>% filter(state == "Kansas" & year_qtr >= 2012 & year_qtr < 2013) 
```


### Using the Synthetic Controls Method
We will begin by running the synthetic controls method on GDP per capita, wages, and the number of establishments. To run the vanilla synthetic controls method using `augsynth`, set `progfunc` to `None` and `scm` to `TRUE`.

#### Single outcomes
First, we will examine each outcome variable separately, beginning with log GDP per capita `lngdpcapita`.

```{r lngdpcapita_syn}
syn_lngdpcapita <- augsynth(lngdpcapita ~ treated, fips, year_qtr, kansas, progfunc="None", scm=T)
summary(syn_lngdpcapita)
```

```{r lngdpcapita_syn_plot}
plot(syn_lngdpcapita)
```

Next we will examine the log total wages per capita. Since this column doesn't already exist in the dataframe, we will create a `lntotalwagescapita` column.
```{r lntotalwagescapita_syn}
kansas$lntotalwagescapita <- log(kansas$totalwagescapita)
syn_lntotalwagescapita <- augsynth(lntotalwagescapita ~ treated, fips, year_qtr, kansas, progfunc="None", scm=T)
summary(syn_lntotalwagescapita)
```

```{r lntotalwagescapita_syn_plot}
plot(syn_lntotalwagescapita)
```

Lastly, we will examine the number of establishments per capita, `estabscapita`.
```{r estabscapita_syn}
syn_estabscapita <- augsynth(estabscapita ~ treated, fips, year_qtr, kansas, progfunc="None", scm=T)
summary(syn_estabscapita)
```

```{r single_estabscapita_syn_plot}
plot(syn_estabscapita)
```

#### Multiple outcomes
Now we will combine our outcome variables into one study. To add more outcome variables, we add them to the LHS of the formula. 
```{r multi_outcome_syn}
syn_multi <- augsynth(lngdpcapita + lntotalwagescapita + estabscapita ~ treated, fips, year_qtr, kansas, progfunc="None", scm=T)

summary(syn_multi, inf = T, inf_type = "conformal", grid_size = 3)
```

```{r multi_outcome_syn_plot}
plot(syn_multi, inf = T, inf_type = "conformal", grid_size = 3)
```


### Using the Augmented Synthetic Controls Method
We will now repeat the study using the Augmented Synthetic Controls Method with ridge regression. In ASCM, we first fit the SCM weights, then combine it with a ridge regression, thus setting `progfunc="Ridge", scm=T`.

```{r lngdpcapita_asyn}
asyn_lngdpcapita <- augsynth(lngdpcapita ~ treated, fips, year_qtr, kansas, progfunc="Ridge", scm=T)
summary(asyn_lngdpcapita)
```

```{r lngdpcapita_asyn_plot}
plot(asyn_lngdpcapita)
```

```{r lntotalwagescapita_asyn}
asyn_lntotalwagescapita <- augsynth(lntotalwagescapita ~ treated, fips, year_qtr, kansas, progfunc="Ridge", scm=T)
summary(asyn_lntotalwagescapita)
```

```{r lntotalwagescapita_asyn_plot}
plot(asyn_lntotalwagescapita)
```

```{r estabscapita_asyn}
asyn_estabscapita <- augsynth(estabscapita ~ treated, fips, year_qtr, kansas, progfunc="Ridge", scm=T)
summary(asyn_estabscapita)
```

```{r single_estabscapita_asyn_plot}
plot(asyn_estabscapita)
```

#### Multiple outcomes
Now we will combine our outcome variables into one study. To add more outcome variables, we add them to the LHS of the formula. 
```{r multi_outcome_asyn}
asyn_multi <- augsynth(lngdpcapita + lntotalwagescapita + estabscapita ~ treated, 
                       fips, year_qtr, kansas, progfunc="Ridge", scm=T, lambda = 1e-4)
summary(asyn_multi, inf = T, inf_type = "conformal", grid_size = 3)
```

```{r multi_outcome_asyn_plot}
plot(asyn_multi, inf = T, inf_type = "conformal", grid_size = 3)
```