---
title: "scpi"
author: "Stefanie Meliss"
date: '`r Sys.Date()`'
output:
  html_document:
    toc: true
    toc_depth: 4
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
################################################################################
## SCPI R Package
## R-file for Empirical Illustration - Single Treated Unit
## Authors: Matias D. Cattaneo, Yingjie Feng, Filippo Palomba, and Rocio Titiunik
################################################################################
### Clear R environment
rm(list = ls(all = TRUE))

### Install R library
# install.packages('scpi')

### Load SCPI package
library(scpi)
library(ggplot2)
library(latex2exp)

set.seed(8894)
```

  
Code from [scpi: Uncertainty Quantification for Synthetic  Control Methods by Cattaneo et al. (2024)](https://nppackages.github.io/references/Cattaneo-Feng-Palomba-Titiunik_2024_JSS.pdf), also see package reference manual [here](https://cran.r-project.org/web/packages/scpi/scpi.pdf).  
  
The package implements estimation, inference procedures, and produces plots for Synthetic Control (SC) methods using least squares, lasso, ridge, or simplex-type constraints. Uncertainty is quantified using prediction intervals according to Cattaneo, Feng, and Titiunik (2021).  
  
Included functions are: *scdata* and *scdataMulti* for data preparation, *scest* for point estimation, *scpi*
for inference procedures, and *scplot* and *scplotMulti* for plots. print() and summary() methods are available for scest and scpi.  
  
# R-file for Empirical Illustration - Single Treated Unit  
  
source: https://github.com/nppackages/scpi/blob/main/R/scpi_illustration.R  
  
## Data preparation  
  
The function scdata() in scpi prepares the data for point estimation/prediction purposes. This function takes as input an object of class DataFrame and outputs an object of class scpi_data containing the matrices A, B, C described above, and a matrix of post-treatment predictors P = (pT0+1, · · · , pT0+T1)′. The user must provide a variable containing a unit identifier (id.var), a time variable (time.var), an outcome variable (outcome.var), the features to be matched (features), the treated unit (unit.tr), the control units (unit.co), the pre-treatment periods (period.pre), and the post-treatment periods (period.post). These options completely specify A, B, and P.  
  
The user can also control the form of R in (1) or, equivalently, the form of C, through the options cov.adj and constant. The former option allows the user to flexibly specify covariate adjustment feature by feature, while the latter option introduces a column vector of ones of size T0 · M in C. If M = 1, this is a simple constant term, but if M ≥ 2 it corresponds to an intercept which is common across features.  The use of the options cov.adj and constant is best explained through some examples.  
  
```{r}
###############################################################################
# SINGLE TREATED UNIT
###############################################################################

### Load data
url <- "https://raw.githubusercontent.com/nppackages/scpi/main/R/scpi_germany.csv"
data <- read.csv(url)

####################################
### Set options for data preparation
id.var      <- "country"                              # ID variable
time.var    <- "year"                                 # Time variable
period.pre  <- seq(from = 1960, to = 1990, by = 1)    # Pre-treatment period
period.post <- (1991:2003)                            # Post-treatment period
unit.tr     <- "West Germany"                         # Treated unit (in terms of id.var)
unit.co     <- setdiff(unique(data$country), unit.tr) # Donors pool
outcome.var <- "gdp"                                  # Outcome variable
cov.adj     <- NULL                                   # Covariates for adjustment
features    <- NULL                                   # No features other than outcome
constant    <- FALSE                                  # No constant term
cointegrated.data <- TRUE                             # Belief that the data are cointegrated
```

### Covariates for adjustment  
  
In the context of Synthetic Control Method (SCM) analysis, features refer to the variables or characteristics used to construct the synthetic control. These variables can include economic indicators, demographic data, or any other relevant metrics that help in creating a comparable control group for the treated unit.

Feature-specific covariate adjustment involves adjusting these features to account for differences between the treated unit and the control group. This adjustment helps to ensure that the synthetic control more accurately represents what would have happened to the treated unit in the absence of the intervention. By doing so, researchers can more reliably attribute any observed differences in outcomes to the intervention itself, rather than to pre-existing differences between the treated unit and the control group.

For example, if you are analysing the impact of a new policy on unemployment rates in a specific region, you would select features such as GDP, population density, and historical unemployment rates. Feature-specific covariate adjustment would then involve making adjustments to these features to align the synthetic control region as closely as possible with the treated region, ensuring that any observed changes in unemployment rates can be attributed to the policy rather than other factors.

In summary, features are the variables used to create the synthetic control, and feature-specific covariate adjustment helps to refine these variables to make the comparison between the treated unit and the synthetic control more accurate. This process enhances the validity of the SCM analysis.

As a means of covariate adjustment in Synthetic Control Method (SCM) analysis, you can include feature-specific constant terms or time trends.

Feature-specific constant terms involve adding a constant value to the features to account for any baseline differences between the treated unit and the control group. This adjustment helps to align the starting levels of the features, making the synthetic control more comparable to the treated unit.

Time trends, on the other hand, involve adjusting the features to account for trends over time. This means you can include terms that capture the trajectory of the features over the period of analysis. By doing so, you can control for any systematic changes that occur over time, ensuring that the synthetic control reflects not just the levels but also the trends in the treated unit's features.  

In summary, including feature-specific constant terms and time trends in your SCM analysis helps to adjust for baseline differences and temporal dynamics, respectively. This ensures a more accurate and reliable comparison between the treated unit and the synthetic control, enhancing the robustness of your findings.

### Non-stationary (cointegrated) data  
  
Non-stationary (cointegrated) data refers to a set of time series variables that are individually non-stationary, meaning their statistical properties such as mean and variance change over time. However, these variables share a common stochastic trend. When combined in a certain way, their linear combination is stationary, meaning the combined series has constant statistical properties over time.

In simpler terms, while the individual series may wander and show trends or patterns over time, there exists a relationship between them that keeps their movements in sync in the long run. This concept is crucial in time series analysis, particularly in econometrics, as it allows for meaningful modelling and forecasting of non-stationary data by taking advantage of their long-term equilibrium relationship.  
  
Cointegrated data allows the user to model the belief that variables A and B form a cointegrated system. This means that although A and B are individually non-stationary, they move together in such a way that their relationship remains stable over time.

In practice, when dealing with the pseudo-true residuals ( u ), you use the first differences of B rather than the levels. This approach helps to address the non-stationarity of B while preserving the cointegrated relationship between A and B. By differencing B, you transform it into a stationary series, making it easier to model and analyse. This method ensures that the long-term equilibrium relationship between A and B is maintained, even though their individual levels may exhibit trends or other non-stationary behaviour.  
  
```{r}
####################################
### Data preparation
df  <-   scdata(df = data, id.var = id.var, time.var = time.var, outcome.var = outcome.var,
                period.pre = period.pre, period.post = period.post,
                unit.tr = unit.tr, unit.co = unit.co, cov.adj = cov.adj, features = features,
                constant = constant, cointegrated.data = cointegrated.data)
# summarise data
print(df)
summary(df)
```

The command returns an object of class 'scdata' containing the following

A	    a matrix containing pre-treatment features of the treated unit.  

B	    a matrix containing pre-treatment features of the control units.  

C	    a matrix containing covariates for adjustment.  

P	    a matrix whose rows are the vectors used to predict the out-of-sample series for the synthetic unit.  

Y.pre	    a matrix containing the pre-treatment outcome of the treated unit.  

Y.post    a matrix containing the post-treatment outcome of the treated unit.  

Y.donors  a matrix containing the pre-treatment outcome of the control units.  


## Point estimation

```{r}
####################################
### SC - point estimation with simplex
est.si  <- scest(data = df, w.constr = list(name="simplex"))
# Use print or summary methods to check results
print(est.si)
summary(est.si)
est.si2 <- scest(data = df, w.constr = list(p = "L1", dir = "==", Q = 1, lb = 0))
summary(est.si2)
```
  
#### Plot results  
  
```{r}
####################################
### SC - plot results
scplot(result = est.si, fig.path = ".",
       fig.name = "germany_est", fig.format = "png", plot.range = (1960:2003),
       label.xy = list(x.lab = "Year", y.lab = "GDP per capita (thousand US dollars)"),
       event.label = list(lab = "Reunification", height = 10))

```

```{r}
####################################
### SC - point estimation with lasso
est.lasso <- scest(data = df, w.constr = list(name="lasso"))
summary(est.lasso)
est.lasso2 <- scest(data = df, w.constr = list(p = "L1", dir = "<=", Q = 1, lb = -Inf),
                    solver="OSQP")
summary(est.lasso2)

```

```{r}
####################################
### SC - point estimation with ridge
est.ridge <- scest(data = df, w.constr = list(name="ridge"))
summary(est.ridge)
Qest <- est.ridge$est.results$w.constr$Q
est.ridge2 <- scest(data = df, w.constr = list(p = "L2", dir = "<=", Q = Qest, lb = -Inf))
summary(est.ridge2)
```
```{r}
####################################
### SC - point estimation with L1-L2
est.l1l2 <- scest(data = df, w.constr = list(name="L1-L2"))
summary(est.l1l2)
est.l1l2.2 <- scest(data = df, w.constr = list(p = "L1-L2", dir = "==/<=", Q = 1, 
                                             Q2 = Qest, lb = 0))
summary(est.l1l2.2)

```

```{r}
####################################
### SC - point estimation with least squares
est.ls <- scest(data = df, w.constr = list(name="ols"))
summary(est.ls)
est.ls2 <- scest(data = df, w.constr = list(p = "no norm", dir = NULL, Q = NULL, lb = -Inf))
summary(est.ls2)
```

## Inference procedures  
  
```{r}
####################################
## Set options for inference
u.alpha  <- 0.05                         # Confidence level (in-sample uncertainty)
e.alpha  <- 0.05                         # Confidence level (out-of-sample uncertainty)
rho      <- NULL                         # Regularization parameter (if NULL it is estimated)
rho.max  <- 0.2                          # Maximum value attainable by rho
sims     <- 200                          # Number of simulations
u.order  <- 1                            # Degree of polynomial in B and C when modelling u
u.lags   <- 0                            # Lags of B to be used when modelling u
u.sigma  <- "HC1"                        # Estimator for the variance-covariance of u
u.missp  <- T                            # If TRUE then the model is treated as misspecified
e.lags   <- 0                            # Degree of polynomial in B and C when modelling e
e.order  <- 1                            # Lags of B to be used when modelling e
e.method <- "gaussian"                   # Estimation method for out-of-sample uncertainty
cores    <- 1                            # Number of cores to be used by scpi
w.constr <- list(name = "simplex")       # Simplex-type constraint set

set.seed(8894)
pi.si   <- scpi(data = df,u.order = u.order, u.lags = u.lags, u.sigma = u.sigma, 
                u.missp = u.missp, sims = sims, e.order = e.order, e.lags = e.lags,
                e.method = e.method, cores = cores, w.constr = w.constr, u.alpha = u.alpha,
                e.alpha = e.alpha, rho = rho, rho.max = rho.max) 

# Use print or summary methods to check results
print(pi.si)
summary(pi.si)
```

#### Plot results
```{r}
####################################
### SC - plot results
scplot(result = pi.si, fig.path = ".",
       fig.name = "germany_unc", fig.format = "png", plot.range = (1960:2003),
       label.xy = list(x.lab = "Year", y.lab = "GDP per capita (thousand US dollars)"),
       x.ticks = NULL, e.out = T, event.label = list(lab = "Reunification", height = 10))
```

## Sensitivity analysis

```{r}
########################################################
# Sensitivity Analysis for 1997 using subgaussian bounds
########################################################
set.seed(8894)
res.si  <- scpi(data = df, sims = sims, e.method = "gaussian", e.order = e.order, e.lags = e.lags,
                u.order = u.order, u.lags = u.lags, u.sigma = u.sigma, u.missp = u.missp,
                cores = cores, w.constr = list(name = "simplex"), lgapp = "linear") 

e.alpha <- 0.05  # default level in scpi
sens <- c(0.25, 0.5, 1, 1.5, 2)
time <- c(1997)
emean <- res.si$inference.results$e.mean
esig <- sqrt(res.si$inference.results$e.var)
sc.l.0 <- res.si$inference.results$CI.in.sample[,1,drop = F]
sc.r.0 <- res.si$inference.results$CI.in.sample[,2,drop = F]
y <- res.si$data$Y.post

for (l in 1:length(time)) {
  ssc.l.1 <- ssc.r.1 <- c()
  e.mean <- emean[time[l]-1990]
  sig <- esig[time[l]-1990]
  sig.seq <- sens * sig

  for (s in 1:length(sig.seq)) {
    eps  <- sqrt(-log(e.alpha/2)*2*(sig.seq[s]^2))
    ssc.l.1[s] <- sc.l.0[time[l]-1990] + e.mean - eps
    ssc.r.1[s] <- sc.r.0[time[l]-1990] + e.mean + eps
  }

  sen.dat <- data.frame(t=c(1:5), lb1=ssc.l.1, ub1=ssc.r.1,
                        lb=rep(sc.l.0[time[l]-1990], 5),
                        ub=rep(sc.r.0[time[l]-1990], 5),
                        lab=as.factor(sens))
  plot <- ggplot() + theme_bw() +
    theme(panel.grid.major = element_blank(),
          panel.grid.minor = element_blank()) +
    labs(x="sd. of e", y="GDP per capita (thousand US dollars)")
  plot <- plot + geom_errorbar(data=sen.dat, aes(x=lab, ymin=lb1, ymax=ub1),
                               col="maroon", width=0.2, linetype=5) +
    geom_errorbar(data=sen.dat, aes(x=lab, ymin=lb, ymax=ub),
                  col="blue", width=0.2, linetype=1) +
    geom_hline(yintercept = y[time[l]-1990], linetype=1, linewidth=0.3, alpha=0.8) +
    annotate("text", x=5.4, y=y[time[l]-1990]-.1,label="Y(1)", size=3.5) +
    scale_x_discrete(labels=c(parse(text=TeX("$0.25\\hat{\\sigma}$")),
                              parse(text=TeX("$0.5\\hat{\\sigma}$")),
                              parse(text=TeX("$\\hat{\\sigma}$")),
                              parse(text=TeX("$1.5\\hat{\\sigma}$")),
                              parse(text=TeX("$2\\hat{\\sigma}$"))))
  print(plot)
}

```
  
## Other data preparation examples  
  
```{r}
################################################################################
### Other examples of data preparation

## multiple features
df  <-   scdata(df = data, id.var = id.var, time.var = time.var, outcome.var = outcome.var,
                period.pre = period.pre, period.post = period.post,
                unit.tr = unit.tr, unit.co = unit.co, cov.adj = cov.adj, features = c("gdp", "trade"),
                constant = constant, cointegrated.data = cointegrated.data)
# summarise data
summary(df)

## multiple features and feature-specific covariate adjustment
df  <-   scdata(df = data, id.var = id.var, time.var = time.var, outcome.var = outcome.var,
                period.pre = period.pre, period.post = period.post,
                unit.tr = unit.tr, unit.co = unit.co, features = c("gdp", "trade"), 
                cov.adj = list('gdp' = c("constant","trend"), 'trade' = c("constant")),
                constant = constant, cointegrated.data = cointegrated.data)
# summarise data
summary(df)

## multiple features and feature-specific covariate adjustment (just for one of the two features)
df  <-   scdata(df = data, id.var = id.var, time.var = time.var, outcome.var = outcome.var,
                period.pre = period.pre, period.post = period.post,
                unit.tr = unit.tr, unit.co = unit.co, features = c("gdp", "trade"), 
                cov.adj = list('gdp' = c("constant","trend"), 'trade' = c()),
                constant = constant, cointegrated.data = cointegrated.data)
# summarise data
summary(df)
```

```{r}
# design matrices
df$A # x1
df$B # x0
df$C # covariates for adjustment
df$P # 

df$Y.pre # z1
df$Y.post # y1
df$Y.donors # z0
```


```{r}
## Specifying features for different pre-treatment periods or just use pre-treatment averages

# I) we want to include "trade" just for some selected periods, i.e., 1960, 1970, 1980, 1990
tradeUnselPeriods <- setdiff(period.pre, seq(1960, 1990, by = 10))  # period NOT to be included
data$tradeAux <- data$trade                                         # copy of the feature of interest
data$tradeAux[data$year %in% tradeUnselPeriods] <- NA

df <- scdata(df = data, id.var = id.var, time.var = time.var, outcome.var = outcome.var,
             period.pre = period.pre, period.post = period.post,
             unit.tr = unit.tr, unit.co = unit.co, features = c("gdp", "tradeAux"), 
             cov.adj = list('gdp' = c("constant","trend"), 'tradeAux' = c("constant")),
             constant = constant, cointegrated.data = cointegrated.data)

df$B
```

```{r}
# II) we want to include just the pre-treatment average of "infrate"
dataAux <- subset(data, year <= 1990) # select just pre-treatment periods
dataAuxAgg <- aggregate(dataAux$infrate, by = list(dataAux$country), FUN = mean, na.rm = TRUE) # take average
names(dataAuxAgg) <- c("country", "infrateAvg")
data <- merge(data, dataAuxAgg, by = c("country")) # merge back
data$infrateAvg[data$year != 1990] <- NA # any period would work, 1990 is arbitrary

df <- scdata(df = data, id.var = id.var, time.var = time.var, outcome.var = outcome.var,
             period.pre = period.pre, period.post = period.post,
             unit.tr = unit.tr, unit.co = unit.co, features = c("gdp", "infrateAvg"),
             cov.adj = list("gdp" = c("constant", "trend")),
             constant = constant, cointegrated.data = cointegrated.data)

df$B
```
  
# R-file for Empirical Illustration Plot - Single Treated Unit  
  
source: https://github.com/nppackages/scpi/blob/main/R/scpi_illustration_plot.R  

 
```{r}
################################################################################
## SCPI R Package
## R-file for Empirical Illustration - Single Treated Unit
## Authors: Matias D. Cattaneo, Yingjie Feng, Filippo Palomba, and Rocio Titiunik  
################################################################################
### Clear R environment
rm(list=ls(all=TRUE))

### Install R library
#install.packages('scpi')

### Load packages
library(scpi)
library(ggplot2)

set.seed(8894)
theme_set(theme_bw())

```
  
## Data preparation  
  
```{r}

##############################################################################
# SINGLE TREATED UNIT
##############################################################################

### Load data
url <- "https://raw.githubusercontent.com/nppackages/scpi/main/R/scpi_germany.csv"
data <- read.csv(url)

####################################
### Set options for data preparation
id.var      <- "country"                              # ID variable
time.var    <- "year"                                 # Time variable
period.pre  <- seq(from = 1960, to = 1990, by = 1)    # Pre-treatment period
period.post <- (1991:2003)                            # Post-treatment period
unit.tr     <- "West Germany"                         # Treated unit (in terms of id.var)
unit.co     <- setdiff(unique(data$country), unit.tr) # Donors pool
outcome.var <- "gdp"                                  # Outcome variable
cov.adj     <- NULL                                   # Covariates for adjustment
features    <- NULL                                   # No features other than outcome
constant    <- FALSE                                  # No constant term
report.missing <- FALSE                               # To check where missing values are
cointegrated.data <- TRUE                             # Belief that the data are cointegrated

```

```{r}
####################################
### Data preparation
df  <-   scdata(df = data, id.var = id.var, time.var = time.var, outcome.var = outcome.var,
                period.pre = period.pre, period.post = period.post,
                unit.tr = unit.tr, unit.co = unit.co, cov.adj = cov.adj, features = features,
                constant = constant, cointegrated.data = cointegrated.data)


```
  
## Point estimation
  
```{r}
####################################
## Set options for inference
u.alpha  <- 0.05                         # Confidence level (in-sample uncertainty)
e.alpha  <- 0.05                         # Confidence level (out-of-sample uncertainty)
rho      <- NULL                         # Regularization parameter (if NULL it is estimated)
rho.max  <- 1                            # Maximum value attainable by rho
sims     <- 200                          # Number of simulations
V        <- NULL                         # Weighting matrix (if NULL it is the identity matrix)
u.order  <- 1                            # Degree of polynomial in B and C when modelling u
u.lags   <- 0                            # Lags of B to be used when modelling u
u.sigma  <- "HC1"                        # Estimator for the variance-covariance of u
u.missp  <- T                            # If TRUE then the model is treated as misspecified
e.lags   <- 0                            # Degree of polynomial in B and C when modelling e
e.order  <- 1                            # Lags of B to be used when modelling e
e.method <- "gaussian"                   # Estimation method for out-of-sample uncertainty
cores    <- 1                            # Number of cores to be used by scpi
w.constr <- list(name = "simplex")       # Simplex-type constraint set

set.seed(8894)
result  <- scpi(data = df,u.order = u.order, u.lags = u.lags, u.sigma = u.sigma, 
                u.missp = u.missp, sims = sims, e.order = e.order, e.lags = e.lags,
                e.method = e.method, cores = cores, w.constr = w.constr, u.alpha = u.alpha,
                e.alpha = e.alpha, rho = rho, rho.max = rho.max) 

```
  
## Plot results

```{r}
####################################
### SC - plot results
scplot(result = result, fig.path = ".",
       fig.name = "germany_unc", fig.format = "png", plot.range = (1960:2003),
       label.xy = list(x.lab = "Year", y.lab = "GDP per capita (thousand US dollars)"),
       x.ticks = NULL, e.out = T, event.label = list(lab = "Reunification", height = 10))

```
  
## Reproduce plot manually
  
```{r}
####################################
### SC - manually reproduce plot
# Store data on treated unit, synthetic unit, and prediction bars
y.fit <- rbind(result$est.results$Y.pre.fit, result$est.results$Y.post.fit)
y.act <- rbind(result$data$Y.pre, result$data$Y.post)

sc.l  <- result$inference.results$CI.all.gaussian[, 1, drop = FALSE]
sc.r  <- result$inference.results$CI.all.gaussian[, 2, drop = FALSE]

# Store other specifics
period.pre  <- result$data$specs$period.pre
period.post <- result$data$specs$period.post
T0          <- period.pre[length(period.pre)] # intercept
plot.range  <- c(period.pre, period.post)

# Actual data
dat    <- data.frame(t     = c(period.pre, period.post),
                     Y.act = c(y.act),
                     sname = "Treated")

# Fill with NAs Y.fit and confidence bounds where missing
Y.fit.na  <- matrix(NA, nrow = length(c(period.pre, period.post)))
sc.l.na   <- matrix(NA, nrow = length(c(period.pre, period.post)))
sc.r.na   <- matrix(NA, nrow = length(c(period.pre, period.post)))

names <- strsplit(rownames(y.fit), "\\.")
not.missing.plot <- c(period.pre,period.post) %in% unlist(lapply(names, "[[", 2))
names <- strsplit(rownames(sc.l), "\\.")
not.missing.ci   <- c(period.pre,period.post) %in% unlist(lapply(names, "[[", 2))

Y.fit.na[not.missing.plot, 1] <- y.fit
sc.l.na[not.missing.ci, 1]    <- sc.l
sc.r.na[not.missing.ci, 1]    <- sc.r

# Synthetic unit data
dat.sc <- data.frame(t        = c(period.pre, period.post),
                     Y.sc     = Y.fit.na,
                     lb       = c(sc.l.na), ub = c(sc.r.na),
                     sname    = "SC Unit")

# Set ticks, event label and merge
x.ticks <- c(seq(plot.range[1], plot.range[length(plot.range)], length.out = 5), T0)
x.ticks <- round(unique(x.ticks))


event.lab <- paste("\n", "Reunification", sep = "")
event.lab.height <- 10

dat.plot    <- subset(dat,    t %in% plot.range)
dat.sc.plot <- subset(dat.sc, t %in% plot.range)

plotdf <- dplyr::left_join(dat.plot, dat.sc.plot, by = 't')

## Plot specs
plot <- ggplot() + theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1)) +
  labs(x = "Year", y = "GDP per capita (thousand US dollars)") +
  theme(legend.position = "bottom", legend.box = "horizontal", legend.title = element_blank(),
        legend.background = element_rect(fill = "white", color = "black"))

## Add Series to plot
plot <- plot + 
  geom_line( data = plotdf, aes(x = t, y = Y.act, colour = sname.x), linetype = 'solid') +
  geom_point(data = plotdf, aes(x = t, y = Y.act, colour = sname.x), shape = 1) +
  geom_line( data = plotdf, aes(x = t, y = Y.sc,  colour = sname.y), linetype = 'dashed') +
  geom_point(data = plotdf, aes(x = t, y = Y.sc,  colour = sname.y), shape = 19) +
  geom_vline(xintercept = T0, linetype = "dashed") +
  geom_text(aes(x = T0, label = event.lab, y = event.lab.height), angle = 90, size = 4) +
  scale_x_continuous(breaks = x.ticks) + 
  scale_color_manual(name = "", values = c("mediumblue", "grey46"),
                     labels = c("Synthetic Control", "Treated"),
                     guide = guide_legend(override.aes = list(
                       linetype = c('dashed','solid'), shape = c(19, 1))))

## Add confidence bars and plot
plot + geom_errorbar(data = plotdf,
                    aes(x = t, ymin = lb, ymax = ub, colour = sname.y),
                    width = 0.5, linetype = 1) + ggtitle("In and Out of Sample Uncertainty")
```




# R-file for Empirical Illustration - Multiple Treated Units

source: https://github.com/nppackages/scpi/blob/main/R/scpi_illustration-multi.R  

```{r}
################################################################################
## SCPI R Package
## R-file for Empirical Illustration - Multiple Treated Units
## Authors: Matias D. Cattaneo, Yingjie Feng, Filippo Palomba, and Rocio Titiunik
################################################################################
### Clear R environment
rm(list = ls(all = TRUE))

### Install R library
#install.packages("scpi")

### Load SCPI package
library(scpi)

set.seed(8894)

###############################################################################
# MULTIPLE TREATED UNITS
###############################################################################

### Load data
data <- scpi_germany

# Create a second placebo treated unit
data$treatment <- 0
data[(data$country == "West Germany" & data$year >= 1991), "treatment"] <- 1
data[(data$country == "Italy" & data$year >= 1992), "treatment"] <- 1

```

  
## Unit-time treatment effect  

### Data preparation
  

```{r}
df <- scdataMulti(data, id.var = "country", outcome.var = "gdp",
                  treatment.var = "treatment", time.var = "year", constant = TRUE,
                  cointegrated.data = TRUE, features = list(c("gdp", "trade")),
                  cov.adj = list(c("constant", "trend")))

```

### Point estimation

```{r}
res <- scest(df, w.constr = list("name" = "simplex"))
```
  
#### Plot results  
  
```{r}
scplotMulti(res)
```


### Inference procedures  
  
```{r}
respi <- scpi(df, w.constr = list("name" = "simplex"), cores = 1, sims = 50,
              e.method = "gaussian")

```

#### Plot results  
  
```{r}
# plot series
scplotMulti(respi, type = "series")

# plot treatment
scplotMulti(respi, type = "series", joint = TRUE)
```

## Average unit treatment effect  

### Data preparation
  
```{r}
df <- scdataMulti(data, id.var = "country", outcome.var = "gdp",
                  treatment.var = "treatment", time.var = "year", constant = TRUE,
                  cointegrated.data = TRUE, features = list(c("gdp", "trade")),
                  cov.adj = list(c("constant", "trend")), effect = "unit")
```

### Point estimation

```{r}
res <- scest(df, w.constr = list("name" = "simplex"))
```
  
#### Plot results  
  
```{r}
scplotMulti(res)
```


### Inference procedures  
  
```{r}
respi <- scpi(df, w.constr = list("name" = "simplex"), cores = 1, sims = 50,
               e.method = "gaussian")

```

#### Plot results  
  
```{r}
# plot series
scplotMulti(respi, type = "series")

# plot treatment
scplotMulti(respi, type = "series", joint = TRUE)
```

## Average unit treatment effect on the treated  

### Data preparation
  
```{r}
df <- scdataMulti(data, id.var = "country", outcome.var = "gdp",
                  treatment.var = "treatment", time.var = "year", constant = TRUE,
                  cointegrated.data = TRUE, features = list(c("gdp", "trade")),
                  cov.adj = list(c("constant", "trend")), effect = "time")
```

### Point estimation

```{r}
res <- scest(df, w.constr = list("name" = "simplex"))
```
  
#### Plot results  
  
```{r}
scplotMulti(res)
```


### Inference procedures  
  
```{r}
respi <- scpi(df, w.constr = list("name" = "simplex"), cores = 1, sims = 50,
               e.method = "gaussian")

```

#### Plot results  
  
```{r}
# plot series
scplotMulti(respi, type = "series")

# plot treatment
scplotMulti(respi, type = "series", joint = TRUE)
```



# R-file for Empirical Illustration Plot - Multiple Treated Units

source: https://github.com/nppackages/scpi/blob/main/R/scpi_illustration_plot-multi.R

```{r}
################################################################################
## SCPI R Package
## R-file for Empirical Illustration - Multiple Treated Units
## Authors: Matias D. Cattaneo, Yingjie Feng, Filippo Palomba, and Rocio Titiunik
################################################################################
### Clear R environment
rm(list=ls(all=TRUE))

### Install R library
#install.packages('scpi')

### Load packages
library(scpi)
library(ggplot2)

set.seed(8894)
theme_set(theme_bw())

```

```{r}
###############################################################################
# MULTIPLE TREATED UNITS
###############################################################################

### Load data
data <- scpi_germany

data$treatment <- 0
data[(data$country == "West Germany" & data$year >= 1991), 'treatment'] <- 1
data[(data$country == "Italy" & data$year >= 1992), 'treatment'] <- 1


###############################################
# unit-time treatment effect
###############################################

df <- scdataMulti(data, id.var = "country", outcome.var = "gdp", 
                  treatment.var = "treatment", time.var = "year", constant = TRUE, 
                  cointegrated.data = T, features = list(c("gdp","trade")),
                  cov.adj = list(c("constant", "trend")))

res.pi <- scpi(df, w.constr = list("name" = "simplex"), cores = 1, sims = 50,
               e.method = "gaussian")

# plot series
scplotMulti(res.pi, type = "series", joint = TRUE, save.data = '__scpi_data')

load('__scpi_data.RData')

plot <- ggplot(toplot) + xlab("Date") + ylab("Outcome") +
  geom_line(aes(x=Time, y=Y, colour=Type)) + 
  geom_point(aes(x=Time, y=Y, colour=Type), size=1.5) + 
  geom_vline(aes(xintercept=Tdate)) +
  facet_wrap(~ID, ncol = 2) + theme(legend.position="bottom") +
  scale_color_manual(name = "", values = c("black", "blue"),
                     labels = c("Treated", "Synthetic Control"))

plot.w1 <- plot + geom_errorbar(data = toplot,
                                aes(x = Time, ymin = lb.gaussian, ymax = ub.gaussian), colour = "blue",
                                width = 0.5, linetype = 1) + ggtitle("In and Out of Sample Uncertainty - Subgaussian Bounds")

plotdf <- subset(toplot, Type == "Synthetic")
plot.w1 + geom_ribbon(data=plotdf, aes(x=Time, ymin=lb.joint, ymax=ub.joint), fill="blue", alpha=0.1)
```

