---
title: "Treated MAT for feasibility evaluations"
author: "Stefanie Meliss"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 4
    toc_float: true
---

Possible options are being evaluated for the following criteria:  

- sufficient amount of pre-treatment observations (min. four/six obs.)  
- timeseries are not too noisy  


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

options(scipen = 999)
# empty work space
rm(list = ls())
gc()

# load libraries
library(kableExtra)
library(dplyr)
library(data.table)
library(ggplot2)

# create function to source code
source_code <- function(root_dir_name = "code", target_repo = "helper_functions", branch = "main", file_name = "file.R") {
  
  # construct URL
  git_url <- paste0("https://raw.githubusercontent.com/stefaniemeliss/", target_repo, "/", branch, "/", file_name)
  
  # attempt to download from github
  tempp_file <- tempfile(fileext = ".R")
  message <- curl::curl_download(git_url, tempp_file, quiet = F)
  
  if(!grepl("Error", message)) {
    
    # if successful, source file
    source(tempp_file)
    remove(tempp_file)
    
  } else { # load local copy of file
    
    # Get the current working directory
    current_dir <- getwd()
    
    # Split the current directory into its components
    dir_components <- strsplit(current_dir, "/")[[1]]
    
    # Identify the root directory dynamically based on the provided root directory name
    root_index <- which(dir_components == root_dir_name)
    if (length(root_index) == 0) {
      stop(paste("Root directory", root_dir_name, "not found in the current path"))
    }
    root_dir <- do.call(file.path, as.list(dir_components[1:root_index]))
    
    # Identify the subdirectory one level below the root and construct its absolute path
    project_repo <- dir_components[root_index + 1]
    dir <- file.path(root_dir, project_repo)
    
    if (target_repo != project_repo) {
      dir <- gsub(project_repo, target_repo, dir) 
    }
    
    # Construct the full file path
    file_path <- file.path(dir, file_name)
    
    # Print the directory and file path for debugging
    print(paste("Directory:", dir))
    print(paste("File path:", file_path))
    
    # Source the file into the parent frame
    source(file_path, local = parent.frame())
  }
}

# source functions
source_code(target_repo = "scm_feasibility", file_name = "functions.R")

# Define the base directory
dir <- get_directory()
dir_data <- file.path(dir, "data")
dir_misc <- file.path(dir, "misc")

# create interim folder
dir.create(path = file.path(getwd(), "interim"))

# get file stem name
file_stem <- get_file_stem()
out_file <- file.path(getwd(), "interim", paste0(file_stem, "_out.csv"))

# copy data #
file.copy(
  file.path(gsub("scm_feasibility", "edu_stats", dir_data), "data_swf.csv"), dir_data, overwrite = T)
file.copy(
  file.path(gsub("scm_feasibility", "edu_stats", dir_data), "data_pupils.csv"), dir_data, overwrite = T)
file.copy(
  file.path(gsub("scm_feasibility", "edu_stats", dir_data), "data_establishments_search.csv"), dir_data, overwrite = T)
file.copy(
  file.path(gsub("scm_feasibility", "edu_stats", dir_data), "data_establishments_groups.csv"), dir_data, overwrite = T)

# load data #

swf <- fread(file.path(dir_data, "data_swf.csv"))
pup <- fread(file.path(dir_data, "data_pupils.csv"))
est <- fread(file.path(dir_data, "data_establishments_search.csv"), na.strings = "")
groups <- fread(file.path(dir_data, "data_establishments_groups.csv"), na.strings = "")

# create lookup 
lookup <- data.frame(time_period = sort(unique(swf$time_period)))
lookup$time_period_str <- insert_slash(lookup$time_period)
lookup$time_period <- as.numeric(substr(lookup$time_period, 0, 4))


# process data establishments #

uid_treated <- 2939

# Dixon schools
dix <- groups %>% filter(group_uid == uid_treated)

# get establishment data from treated schools
list_laestab <- c(dix$laestab)

est_treated <- est %>% 
  filter(laestab %in% list_laestab) %>%
  mutate(school = "treated") %>% 
  as.data.frame()

run_gridsearch <- F

# define variables
dv = "pupil_to_qual_teacher_ratio"
var_teach = "fte_avg_age"
var_pup = "pnpupfsm_e"

vars = c(dv, var_teach, var_pup)

id_cols <- c("laestab", "time_period", "time_period_str")
```

```{r process_school, echo = F, eval=FALSE}
#id_treated = list_laestab[1] # debug

# ---- Dataset creation ----
# Filter School Workforce (SWF) data to create outcome dataset with selected variables
z <- swf %>%
  filter(laestab %in% id_treated) %>%
  select(time_period, laestab, school, !!sym(dv), !!sym(var_teach))


# Filter pupil data to create predictor dataset with selected variables
x <- pup %>% 
  filter(laestab %in% id_treated) %>%
  select(time_period, laestab, !!sym(var_pup))

# Combine outcome and predictor datasets
df <- merge(z, x, all = T, by = c("laestab", "time_period"))

# Remove any rows with missing values
df <- na.omit(df)

# arrange data
df <- df %>%
  arrange(laestab, desc(time_period)) %>%
  mutate(
    # use consistent school name
    school = first(school),
    # apply rolling window average
    across(all_of(vars), ~ zoo::rollapply(.x, width = 2, mean, align = "left", partial = T), .names = "{col}_roll"),
    # add slash and use as string
    time_period_str = insert_slash(time_period),
    # remove the last two digits
    time_period = as.numeric(substr(time_period, 0, 4))
  ) %>% 
  as.data.frame()


# ONLY PROCEED IF THERE IS DATA FROM AT LEAST 4 YEARS AVAILABLE

cat("## ", unique(df$school), "\n\n")

# ---- Spaghetti plot ----

# create data in long format
df_long <- df %>%
  tidyr::pivot_longer(
    cols = matches("pup|fte"),
    names_to = "variable") %>%
  mutate(category = case_match(variable, 
                               "pupil_to_qual_teacher_ratio" ~ "Outcome",
                               "pupil_to_qual_teacher_ratio_roll" ~ "Outcome",
                               "fte_avg_age" ~ "Teacher age",
                               "fte_avg_age_roll" ~ "Teacher age",
                               "pnpupfsm_e" ~ "% pupils FSM",
                               "pnpupfsm_e_roll" ~ "% pupils FSM"
  ),
  computation = case_match(variable, 
                           "pupil_to_qual_teacher_ratio" ~ "Raw",
                           "pupil_to_qual_teacher_ratio_roll" ~ "Rolling average",
                           "fte_avg_age" ~ "Raw",
                           "fte_avg_age_roll" ~ "Rolling average",
                           "pnpupfsm_e" ~ "Raw",
                           "pnpupfsm_e_roll" ~ "Rolling average"
  )
  ) %>%
  group_by(variable) %>%
  mutate(out_iqr = is_outlier_iqr(value)) %>%
  as.data.frame()

df_long$category <- factor(df_long$category, levels = c("Outcome", "% pupils FSM", "Teacher age"))


# plot timeseries for each variable
plt <- ggplot(data = df_long, aes(x = time_period_str, y = value, col = computation, group = interaction(laestab, computation))) +
  geom_line() + 
  geom_point(aes(shape = out_iqr), show.legend = F) +
  facet_wrap(~ category, ncol = 1, strip.position = "top", scales = "free_y") +
  ambition_theme +
  ylab("Reported value") +
  scale_color_manual(values = c("Raw" = blue, "Rolling average" = red)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title.x = element_blank(), 
        legend.title = element_blank())

print(plt)
cat("\n\n")  # Add some space between plots




#### CREATE SUMMARY TABLE ####

# compute timeseries descriptives
ts_desc <- apply(df[, grepl("pup|fte", names(df))], MARGIN = 2, FUN = function(x){psych::describe(x, IQR = T, trim = 1/nrow(df))})

# combine to df
ts_desc <- do.call("rbind",ts_desc)
ts_desc$vars <- NULL

ts_desc$out_sd <- apply(df[, grepl("pup|fte", names(df))], MARGIN = 2, FUN = function(x){sum(is_outlier_3sd(x))})
ts_desc$out_iqr <- apply(df[, grepl("pup|fte", names(df))], MARGIN = 2, FUN = function(x){sum(is_outlier_iqr(x))})

# compute snr
ts_desc$snr <- apply(df[, grepl("pup|fte", names(df))], MARGIN = 2, FUN = function(x){calculate_snr(x, window_size = 3)})

# compute relative standard deviation
ts_desc$rsd <- ts_desc$sd / ts_desc$mean

if (nrow(df) > 1) {
  
  # compute timeseries auto correlation
  max_lag = ceiling(.25 * nrow(df)) # use lags up to about one-quarter of the total number of observations.
  ts_ac <- apply(df[, grepl("pup|fte", names(df))], MARGIN = 2, FUN = function(x){
    tmp <- acf(ts(x), lag.max = max_lag, plot = F)
    return(tmp$acf)
  })
  
  # transpose output
  ts_ac <- as.data.frame(t(ts_ac))
  names(ts_ac) <- paste0("ac_l", 0:(ncol(ts_ac)-1))
  ts_ac$ac_l0 <- NULL
  
  # # compute Augmented Dickey-Fuller (ADF) test
  # ts_adf <- apply(df[, grepl("pup|fte", names(df))], MARGIN = 2, FUN = function(x){
  #   tmp <- tseries::adf.test(ts(x))
  #     return(tmp$p.value)
  #   })
  # ts_adf <- as.data.frame(ts_adf)
  
  # combine all measures
  out <- merge(ts_desc, ts_ac, by = 0)
  # row.names(out) <- out$Row.names
  # out$Row.names <- NULL
  # out <- merge(out, ts_adf, by = 0)
  names(out)[1] <- "Variable"
} else {
  out <- ts_desc
  out$Variable <- row.names(out)
  out <- out %>% relocate(Variable) %>% as.data.frame()
}

# assign new variable names
out$Variable <- factor(out$Variable, 
                       levels = c("pupil_to_qual_teacher_ratio", "pupil_to_qual_teacher_ratio_roll",  
                                  "pnpupfsm_e", "pnpupfsm_e_roll",
                                  "fte_avg_age", "fte_avg_age_roll"),
                       labels = c("Outcome", "Outcome (roll)", 
                                  "FSM", "FSM (roll)",
                                  "Age", "Age (roll)")
                       
)
# sort 
out <- out[order(out$Variable), ]


# print to markdown
kbl(out, caption = paste0("Descriptives of timeseries data"), digits = 2, row.names = F) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), fixed_thead = T) %>% 
  column_spec(c(2, 18), border_right = T) %>%
  column_spec(c(1, 4, 17), bold = T) %>%
  print()
cat("\n")

# save data
if (id_treated == list_laestab[1]) {
  summary <- out %>%
    mutate(school = unique(df$school),
           laestab = unique(df$laestab)) %>%
    relocate(laestab, school) %>%
    as.data.frame()
} else {
  summary <- rbind.all.columns(summary, out %>%
                                 mutate(school = unique(df$school),
                                        laestab = unique(df$laestab)) %>%
                                 relocate(laestab, school) %>%
                                 as.data.frame())
}

```

# School-level data  
  
```{r school, echo=FALSE, results='asis'}

for (id_treated in list_laestab) {
  <<process_school>>
}

# save summary
write.csv(summary, out_file, row.names = F)
summary <- read.csv(out_file)
cat("## Summary \n\n")

# print to markdown
summary <- 
  read.csv(out_file) %>% 
  #filter(Variable %in% c("Outcome (qual)", "FSM (roll)", "Age (roll)")) %>%
  mutate_if(is.numeric, round, digits = 2) %>%
  as.data.frame() 

cat("In total, data from", length(unique(summary$school)), "schools were considered.")
cat("\n\n")
cat(length(unique(summary$school[summary$n >= 4])), "schools had four or more pre-treatment observations.")
cat("\n\n")
# 
summary %>% 
  filter(n >= 4) %>% 
  select(-laestab) %>%
  # mutate(Variable = factor(Variable, levels = c("Outcome (qual)", "FSM (roll)", "Age (roll)"))) %>%
  arrange(school, Variable) %>%
  kbl(caption = paste0("Descriptives of timeseries data"), row.names = F) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), fixed_thead = T) %>% 
  column_spec(c(3, 19), border_right = T) %>%
  column_spec(c(1, 2, 5, 18), bold = T) %>%
  print()
cat("\n\n\n")

```


# MAT-level data  
  
## Data preparation process for synthetic control analysis

##### Longitudinal data filtering

- Keeps only schools with **4 or more years of observations**.


##### MAT-level filtering

- Identifies MATs meeting specific criteria (**more than 1 school in MAT** after school-level filtering, schools only in specified regions).


##### Data aggregation

- Computes MAT-level averages for each academic year.  
- Data from **more than one school** per academic year required.
- Retains only MATs with complete time series data.  
  
  
```{r check_ts, echo = F, eval=FALSE}
# get school-level data of treated MAT
df_s <- df %>% filter(laestab %in% list_laestab)

# create data in long format - school level
df_s <- df_s %>%
  tidyr::pivot_longer(
    cols = c(pupil_to_qual_teacher_ratio, fte_avg_age, pnpupfsm_e),
    names_to = "variable") %>%
  mutate(category = case_match(variable, 
                               "pupil_to_qual_teacher_ratio" ~ "Outcome",
                               "fte_avg_age" ~ "Teacher age",
                               "pnpupfsm_e" ~ "% pupils FSM")) %>%
  as.data.frame()

df_s$category <- factor(df_s$category, levels = c("Outcome", "Teacher age", "% pupils FSM"))

# create data in long format - MAT
df_long <- df_treat %>%
  tidyr::pivot_longer(
    cols = c(pupil_to_qual_teacher_ratio, fte_avg_age, pnpupfsm_e),
    names_to = "variable") %>%
  mutate(category = case_match(variable, 
                               "pupil_to_qual_teacher_ratio" ~ "Outcome",
                               "fte_avg_age" ~ "Teacher age",
                               "pnpupfsm_e" ~ "% pupils FSM"),
         min_schools_per_year = ifelse(n >= 4, "4S/Y", 
                                   ifelse(n == 3, "3S/Y", 
                                          ifelse(n == 2, "2S/Y", NA)))) %>%
  as.data.frame()

df_long$category <- factor(df_long$category, levels = c("Outcome", "Teacher age", "% pupils FSM"))

##### CREATE SPAGHETTI PLOT ####

# plot school-level timeseries for each variable
plt <- ggplot(data = df_s, aes(x = time_period, y = value, group = establishmentname, col = establishmentname, fill = establishmentname, shape = phaseofeducation_name)) +
  geom_line(linewidth = 1) + 
  geom_point(col = black) +
  facet_wrap(~ category, ncol = 1, strip.position = "top", scales = "free_y") +
  ambition_theme +
  ylab("Reported value") + xlab("Academic year") +
  scale_x_continuous(breaks = lookup$time_period, labels = lookup$time_period_str) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.title = element_blank(),
        legend.text = element_text(size = 8),
        legend.position = "bottom", legend.box="vertical",
        legend.spacing = unit(0.2, "cm"), 
        legend.margin = margin(t = 0.1, r = 0.1, b = 0.1, l = 0.1, unit = "cm")) + 
  scale_colour_manual(values = as.vector(pals::glasbey(length(list_laestab)))) +
  scale_fill_manual(values = as.vector(pals::glasbey(length(list_laestab)))) +
  scale_shape_manual(values = c(21:24)) +
  guides(col = guide_legend(nrow = nrow_legend, byrow = T))
print(plt)
cat("\n\n")  # Add some space between plots


# plot MAT AVERAGE timeseries for each variable
# note: school-level data is added to the plot but in transparent (col = NA)
# this keeps the scales constant
plt <- ggplot(data = df_long, aes(x = time_period, y = value, group = group_uid)) +
  geom_point(data = df_s, aes(x = time_period, y = value), col = NA) + 
  #geom_vline(xintercept = period.pre[length(period.pre)]+0.5, linetype = "dotted") +
  geom_line(col = black, linewidth = .8) + 
  geom_point(aes(colour = min_schools_per_year), size = 3) +
  facet_wrap(~ category, ncol = 1, strip.position = "top", scales = "free_y") +
  ambition_theme +
  ylab("Reported value") + xlab("Academic year") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        #legend.title = element_blank(),
        legend.position = "bottom") +
  scale_x_continuous(breaks = lookup$time_period, labels = lookup$time_period_str) +
  scale_color_manual(values = ambition_palette_bright)

suppressWarnings(print(plt))
cat("\n\n")  # Add some space between plots

# plot MAT AVERAGE timeseries for each variable
# add school-level data visibly as POINTS
plt <- ggplot() + 
  geom_point(data = df_s, aes(x = time_period, y = value, fill = establishmentname, shape = phaseofeducation_name, colour = establishmentname)) + 
  geom_line(data = df_long, aes(x = time_period, y = value, group = group_uid), linewidth = 1) + 
  facet_wrap(~ category, ncol = 1, strip.position = "top", scales = "free_y") + 
  ambition_theme + 
  ylab("Reported value") + 
  xlab("Academic year") + 
  scale_x_continuous(breaks = lookup$time_period, labels = lookup$time_period_str) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), 
        legend.title = element_blank(), 
        legend.text = element_text(size = 8), 
        legend.position = "bottom", legend.box="vertical",
        legend.spacing = unit(0.2, "cm"), 
        legend.margin = margin(t = 0.1, r = 0.1, b = 0.1, l = 0.1, unit = "cm")) + 
  scale_fill_manual(values = as.vector(pals::glasbey(length(list_laestab)))) + 
  scale_shape_manual(values = c(21:24)) + 
  scale_colour_manual(values = as.vector(pals::glasbey(length(list_laestab)))) +
  guides(col = guide_legend(nrow = nrow_legend, byrow = T))

print(plt)
cat("\n\n")  # Add some space between plots

# plot MAT AVERAGE timeseries for each variable
# add school-level data visibly as LINES
plt <- ggplot() + 
  geom_line(data = df_s, aes(x = time_period, y = value, colour = establishmentname, group = establishmentname), linewidth = .5) + 
  geom_line(data = df_long, aes(x = time_period, y = value, group = group_uid), linewidth = 1) + 
  facet_wrap(~ category, ncol = 1, strip.position = "top", scales = "free_y") + 
  ambition_theme + 
  ylab("Reported value") + 
  xlab("Academic year") + 
  scale_x_continuous(breaks = lookup$time_period, labels = lookup$time_period_str) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), 
        legend.title = element_blank(), 
        legend.text = element_text(size = 8), 
        legend.position = "bottom", legend.box="vertical",
        legend.spacing = unit(0.2, "cm"), 
        legend.margin = margin(t = 0.1, r = 0.1, b = 0.1, l = 0.1, unit = "cm")) + 
  scale_fill_manual(values = as.vector(pals::glasbey(length(list_laestab)))) + 
  scale_shape_manual(values = c(21:24)) + 
  scale_colour_manual(values = as.vector(pals::glasbey(length(list_laestab)))) +
  guides(col = guide_legend(nrow = nrow_legend, byrow = T))

print(plt)
cat("\n\n")  # Add some space between plots

# print averaged TS to markdown
df_treat %>%
  select(-c(group_uid, status, time_period)) %>%
  relocate(time_period_str) %>%
  relocate(., n, .after = time_period_str) %>%
  rename(., n_schools = n) %>%
  kbl(caption = paste0("MAT timeseries data"), row.names = F, digits = 3) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), fixed_thead = T) %>% 
  print()

# compute timeseries descriptives
ts_desc <- apply(df_treat[, grepl("pup|fte", names(df_treat))], MARGIN = 2, FUN = function(x){psych::describe(x, IQR = T, trim = 1/nrow(df_treat))})

# combine to df
ts_desc <- do.call("rbind",ts_desc)
ts_desc$vars <- NULL

ts_desc$out_sd <- apply(df_treat[, grepl("pup|fte", names(df_treat))], MARGIN = 2, FUN = function(x){sum(is_outlier_3sd(x))})
ts_desc$out_iqr <- apply(df_treat[, grepl("pup|fte", names(df_treat))], MARGIN = 2, FUN = function(x){sum(is_outlier_iqr(x))})

# compute snr
ts_desc$snr <- apply(df[, grepl("pup|fte", names(df))], MARGIN = 2, FUN = function(x){calculate_snr(x, window_size = 3)})

# compute relative standard deviation
ts_desc$rsd <- ts_desc$sd / ts_desc$mean

# compute timeseries auto correlation
max_lag = ceiling(.25 * nrow(df_treat)) # use lags up to about one-quarter of the total number of observations.
ts_ac <- apply(df_treat[, grepl("pup|fte", names(df_treat))], MARGIN = 2, FUN = function(x){
  df_treat <- acf(ts(x), lag.max = max_lag, plot = F)
  return(df_treat$acf)
})

# transpose output
ts_ac <- as.data.frame(t(ts_ac))
names(ts_ac) <- paste0("ac_l", 0:(ncol(ts_ac)-1))
ts_ac$ac_l0 <- NULL

# combine all measures
out <- merge(ts_desc, ts_ac, by = 0)
names(out)[1] <- "Variable"

# assign new variable names
out$Variable <- factor(out$Variable, 
                       levels = c("pupil_to_qual_teacher_ratio",  
                                  "pnpupfsm_e",
                                  "fte_avg_age"),
                       labels = c("Outcome", 
                                  "FSM", 
                                  "Age")
                       
)
# sort 
out <- out[order(out$Variable), ]

# print to markdown
out %>% 
  mutate(Variable = factor(Variable, levels = c("Outcome", "FSM", "Age"))) %>%
  arrange(Variable, sd) %>%
  kbl(caption = paste0("Descriptives of MAT timeseries data"), row.names = F, digits = 3) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), fixed_thead = T) %>% 
  column_spec(c(2, 18), border_right = T) %>%
  column_spec(c(1, 4, 17), bold = T) %>%
  print()
cat("\n")

```

```{r grid_search, echo = F, eval=FALSE}
file_name <- file.path(getwd(), "interim", paste0(file_stem, "_", tolower(phase), ".csv"))

if (run_gridsearch) {
  
  # Define timeseries
  period.avail <- c(2010:2023)
  period.post <- c(2024:2026)
  
  ####################################
  ### define grid
  
  # define different filter option for each phase #
  
  if (phase == "mixed") {
    
    # define options for filtering
    filter.options <- list(
      "NULL",  # This will be interpreted as NULL later
      "! time_period %in% c(201011, 201112)",
      "! laestab %in% c(3802008)",  # This will be interpreted as NULL later
      "! laestab %in% c(3802008) & ! time_period %in% c(201011, 201112)"
    )
    
    # define options for donor pool
    min.years.obs.options = c(4, 6, 8)
    min.schools.per.mat.options = c(2, 3, 4)
    min.schools.per.timeperiod.options = c(2, 3, 4)
    
    regions.options = list(c("Yorkshire and the Humber", "North West"), c("Yorkshire and the Humber"))
    
    exclude.single.phase.options = c(TRUE, FALSE)
    exclude.northwest.options = c(TRUE, FALSE)
    
    filter_phase = c("Not applicable", "16 plus")
    
  } else if (phase == "Secondary") {
    
    # define options for filtering
    filter.options <- list(
      "! time_period %in% c(201011, 201112)",
      "! time_period %in% c(201011, 201112, 201213, 201314)"
    )
    
    # define options for donor pool
    min.years.obs.options = c(4, 6, 8)
    min.schools.per.mat.options = c(2, 3, 4)
    min.schools.per.timeperiod.options = c(2, 3, 4)
    
    regions.options = list(c("Yorkshire and the Humber", "North West"))
    
    exclude.single.phase.options = c(FALSE)
    exclude.northwest.options = c(TRUE, FALSE)
    
    filter_phase = unique(groups$phaseofeducation_name)[!grepl("econdary", unique(groups$phaseofeducation_name))]
  } else if (phase == "Primary") {
    
    # define options for filtering
    filter.options <- list(
      "! time_period %in% c(201011, 201112, 201213, 201314)",
      "! time_period %in% c(201011, 201112, 201213, 201314, 201415)",
      "! laestab %in% c(3802008) & ! time_period %in% c(201011, 201112, 201213, 201314)",
      "! laestab %in% c(3802008) & ! time_period %in% c(201011, 201112, 201213, 201314, 201415)"
    )
    
    # define options for donor pool
    min.years.obs.options = c(4, 6, 8)
    min.schools.per.mat.options = c(2, 3)
    min.schools.per.timeperiod.options = c(2, 3)
    
    filter_phase = unique(groups$phaseofeducation_name)[!grepl("imary", unique(groups$phaseofeducation_name))]
    
  }
  
  
  # Create initial parameter grid
  param_grid <- expand.grid(
    min.years.obs = min.years.obs.options,
    min.schools.per.mat = min.schools.per.mat.options, 
    min.schools.per.timeperiod = min.schools.per.timeperiod.options,
    swf.filter.idx = 1:length(filter.options),
    stringsAsFactors = FALSE
  )
  
  # translate SWF filter #

  # Add the actual filter expressions
  param_grid$swf.filter <- filter.options[param_grid$swf.filter.idx]
  
  # Remove the index column
  param_grid$swf.filter.idx <- NULL
  
  # Add if any schools should be exclude
  param_grid$excl.outlier <- grepl("laestab", param_grid$swf.filter)
  
  # extract any years to be excluded
  param_grid$period.excl <- I(apply(param_grid, 1, function(row) {
    extract_years_from_filter(row["swf.filter"])
  }))
  
  # define period.post and period.avail based on CV filter
  param_grid$period.post <- I(list(period.post))
  param_grid$period.avail <- I(list(period.avail))
  
  # Calculate period.pre based on period.avail, period.post and period.excl.
  param_grid$period.pre <- mapply(calculate_period_pre, param_grid$period.avail, param_grid$period.post, param_grid$period.excl, SIMPLIFY = FALSE)
  
  # create empty df
  df_empty = data.frame(id = integer(1),
                        name = character(1),
                        time_period = numeric(1),
                        time_period_str = character(1),
                        pupil_to_qual_teacher_ratio = numeric(1),
                        fte_avg_age = numeric(1),
                        pnpupfsm_e = numeric(1),
                        n = integer(1),
                        level = character(1)
  )
  
  
  results <- do.call(rbind, lapply(1:nrow(param_grid[, ]), function(i) {
    message(i)
    
    # Extract params
    params <- param_grid[i, ]
    
    # Convert "NULL" string to actual NULL
    swf.filter.param <- if(params$swf.filter == "NULL") NULL else unlist(params$swf.filter)
    params$swf.filter <- unlist(params$swf.filter)
    params$period.excl <- paste(params$period.excl, collapse = ", ")
    params$period.post <- paste(params$period.post, collapse = ", ")
    params$period.avail <- paste(params$period.avail, collapse = ", ")
    params$period.pre <- paste(params$period.pre, collapse = ", ")
    
    # create df_it
    df_it <- merge(params, df_empty)
    
    # run processing with the parameters
    result <- tryCatch({
      process_data_scm_mat(uid_treated = uid_treated, 
                           target_regions = regions, 
                           min_years_obs = params$min.years.obs,
                           min_schools_per_mat = params$min.schools.per.mat,
                           min_schools_per_timeperiod = params$min.schools.per.timeperiod,
                           filter_phase = filter_phase,
                           swf_filter = swf.filter.param)
    }, error = function(e) { 
      return(df_it)
      # return(list(error = paste("Error in process_data_scm_mat:", e$message)))
    })
    
    # Extract information about iteration #
    
    # if there is a MAT timeseries
    if (nrow(result$df_treat) > 0) {
      
      # select relevant columns: MAT
      df_treat <- result$df_treat %>%
        rename(., id = group_uid, name = status) %>%
        select(c(id, name, time_period, time_period_str, 
                 pupil_to_qual_teacher_ratio, fte_avg_age, pnpupfsm_e, n)) %>%
        mutate(level = "MAT")
      
      # select relevant columns: school
      df_s <- result$df %>%
        filter(laestab %in% list_laestab) %>%
        rename(., id = laestab, name = establishmentname) %>%
        select(c(id, name, time_period, time_period_str, 
                 pupil_to_qual_teacher_ratio, fte_avg_age, pnpupfsm_e)) %>%
        mutate(n = NA,
               level = "School")
      
      # combine both
      df_it <- rbind(df_treat, df_s)
      
      # merge with parameter info
      df_it <- merge(params, df_it)
    }
    return(df_it)
  }))
  
  # save results of grid search
  write.csv(results, file = file_name, row.names = F)
  
} else {
  results <- read.csv(file = file_name)
}

```


```{r grid_plot, echo = F, eval=FALSE}
#### process data grid search ####

# copy data for editing
tmp_results <- results

# How was the average timeseries obtained?
tmp_results$min.years.obs <- paste0(tmp_results$min.years.obs, "Y")
tmp_results$min.years.obs <- factor(tmp_results$min.years.obs)
tmp_results$min.schools.per.timeperiod <- paste0(tmp_results$min.schools.per.timeperiod, "S/Y")
tmp_results$min.schools.per.timeperiod <- factor(tmp_results$min.schools.per.timeperiod)
tmp_results$min.schools.per.mat <- paste0(tmp_results$min.schools.per.mat, "S")
tmp_results$min.schools.per.mat <- factor(tmp_results$min.schools.per.mat)
# which years were used in training
# # Function to extract min and max year
# extract_min_max_years <- function(period_string) {
#   years <- as.numeric(unlist(strsplit(period_string, ", ")))
#   min_year <- min(years)
#   max_year <- max(years)
#   return(c(min_year, max_year))
# }
# tmp_results$min_year <- sapply(tmp_results$period.pre, function(x) extract_min_max_years(x)[1])
# tmp_results$max_year <- sapply(tmp_results$period.pre, function(x) extract_min_max_years(x)[2])
# tmp_results$period.pre <- paste0(tmp_results$min_year, ":", tmp_results$max_year)
tmp_results$period.pre <- factor(tmp_results$period.pre)

# did we remove outliers
tmp_results$excl.outlier <- ifelse(tmp_results$excl.outlier, "excl. out", "incl. out")

tmp_results$error <- is.na(tmp_results$pupil_to_qual_teacher_ratio)


#### plot treated timeseries ####
cat("#### All schools \n\n")

plt <- ggplot(data = tmp_results[tmp_results$level == "MAT" & tmp_results$excl.outlier == "incl. out", ], aes(x = time_period, y = pupil_to_qual_teacher_ratio, group = id)) + 
  geom_point(data = tmp_results[tmp_results$level == "School" & tmp_results$excl.outlier == "incl. out", ], aes(x = time_period, y = pupil_to_qual_teacher_ratio, col = name), size = .5) + 
  geom_line(linewidth = .8) +
  ggh4x::facet_grid2(cols = vars(min.years.obs, min.schools.per.timeperiod), rows = vars(period.pre, min.schools.per.mat), strip = ggh4x::strip_nested()) +
  ambition_theme +
  theme(strip.text = element_text(size = 10),
        legend.title = element_blank(),
        legend.text = element_text(size = 8),
        legend.position = "bottom", legend.box="vertical",
        legend.spacing = unit(0.2, "cm"), 
        legend.margin = margin(t = 0.1, r = 0.1, b = 0.1, l = 0.1, unit = "cm"),
        axis.title = element_text(size = 10),
        axis.text = element_text(size = 10),
        # axis.title.x = element_blank(),
        axis.text.x = element_blank()) +
  ylab("Reported value") + xlab("Academic year") +
  scale_colour_manual(values = as.vector(pals::glasbey(length(list_laestab)))) +
  guides(col = guide_legend(nrow = nrow_legend, byrow = T))

print(plt)
cat("\n\n")


if(sum(tmp_results$excl.outlier == "excl. out") > 0) {
  cat("#### Outlier removed \n\n")
  
plt <- ggplot(data = tmp_results[tmp_results$level == "MAT" & tmp_results$excl.outlier == "excl. out", ], aes(x = time_period, y = pupil_to_qual_teacher_ratio, group = id)) + 
  geom_point(data = tmp_results[tmp_results$level == "School" & tmp_results$excl.outlier == "excl. out", ], aes(x = time_period, y = pupil_to_qual_teacher_ratio, col = name), size = .5) + 
  geom_line(linewidth = .8) +
  ggh4x::facet_grid2(cols = vars(min.years.obs, min.schools.per.timeperiod), rows = vars(period.pre, min.schools.per.mat), strip = ggh4x::strip_nested()) +
  ambition_theme +
  theme(strip.text = element_text(size = 10),
        legend.title = element_blank(),
        legend.text = element_text(size = 8),
        legend.position = "bottom", legend.box="vertical",
        legend.spacing = unit(0.2, "cm"), 
        legend.margin = margin(t = 0.1, r = 0.1, b = 0.1, l = 0.1, unit = "cm"),
        axis.title = element_text(size = 10),
        axis.text = element_text(size = 10),
        # axis.title.x = element_blank(),
        axis.text.x = element_blank()) +
  ylab("Reported value") + xlab("Academic year") +
  scale_colour_manual(values = as.vector(pals::glasbey(length(list_laestab)))) +
  guides(col = guide_legend(nrow = nrow_legend, byrow = T))
  
  print(plt)
  cat("\n\n")
  
  
}
```



```{r mat, echo=F, message=FALSE, results='asis', fig.align='center', fig.height=10, fig.width=8, out.width='100%', out.height='100%'}
# define phases to loop through
phases <- c("mixed", "Secondary", "Primary")
headings <- c("Whole MAT", "Secondary schools only", "Primary schools only")
p = 1 # debug
for (p in 1:length(phases)) {
  
  cat("\n\n")
  cat("## ", headings[p], "\n\n")
  
  phase = phases[p]
  
  # define different filter option for each phase #
  
  if (phase == "mixed") {
    swf_filter = NULL
    filter_phase = c("Not applicable", "16 plus")
    nrow_legend = 3
  } else if (phase == "Secondary") {
    swf_filter = "! time_period %in% c(201011, 201112)"
    filter_phase = unique(groups$phaseofeducation_name)[!grepl("econdary", unique(groups$phaseofeducation_name))]
    nrow_legend = 2
  } else if (phase == "Primary") {
    swf_filter = "! time_period %in% c(201011, 201112, 201213, 201314)"
    filter_phase = unique(groups$phaseofeducation_name)[!grepl("imary", unique(groups$phaseofeducation_name))]
    nrow_legend = 1
  }
  
  # Define target regions for filtering the donor pool
  regions <- c("Yorkshire and the Humber", "North West")
  
  # process data at MAT level
  process_data_scm_mat(uid_treated = uid_treated, target_regions = regions, filter_phase = filter_phase,
                       swf_filter = swf_filter, min_years_obs = 4, min_schools_per_mat = 2, min_schools_per_timeperiod = 2)
  
  cat("### Timeseries data \n\n")
  <<check_ts>>
    
    cat("### Different pre-processing settings \n\n")
  
  <<grid_search>>
    <<grid_plot>>
    
}

```



